### 2018-08-16 positional_information_childes.Rmd
Initial information

---
title: Information Content Across Sentence Position in CHILDES
author: Joe Klafka
date: "8/16/2018"
output: html_document
---

```{r setup, include=FALSE}
# load packages
library(knitr)
library(tidyverse)
library(directlabels)
library(childesr) #data
library(SnowballC) #stemmer
library(tidytext)
library(entropy)
library(tidyboot)
library(dplyr)
library(tokenizers)
library(tau)
library(reticulate)
knitr::opts_chunk$set(echo = TRUE)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 16))
```

```{r get_utterances}
prov_utterances <- get_utterances(corpus = "Providence")
```

```{r mutual information}

kid_info <- prov_utterances %>%
          filter(speaker_role == "Target_Child") %>%
          mutate(length = str_count(stem, " ") + 1) %>% 
          filter(length == 5)

```

```{python, engine.path="~/../../usr/local/bin/python3"}

```