### 2018-09-14 wiki_entropy.Rmd
added Wikipedia script and analyses for Georgian, Kyrgyz, Basque and Danish

---
title: Wikipedia Text Mining and Entropy Analysis
author: Josef Klafka
date:
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
---

```{r setup, include=FALSE}
# load packages
library(DBI)
library(knitr)
library(tidyverse)
library(directlabels)
library(dplyr)

library(SnowballC) #stemmer
library(tidytext)
library(entropy)
library(tidyboot)
library(tokenizers)
library(tau)
library(gtools)
library(reticulate)
knitr::opts_chunk$set(echo = TRUE)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 16))
use_python("~/../../Library/Frameworks/Python.framework/Versions/3.7/bin/python3.7")
```

```{python}
import regex as re 
import numpy as np
import pandas as pd

def extract_texts(upper_bound):
	rs = ""
	for i in range(upper_bound):
		if i < 10:
			number = "0" + str(i)
		else:
			number = str(i)
		f = open("~/Documents/Labword/Joe/misc/text/AA/wiki_" + number, 'r')
		text = f.read()
		rs += text
	return rs

def get_sen_dict(text):
	text = re.sub("<doc.+>|</doc>|http:\S+|[-%,;:–'&*#/—“»]|\d+|\(|\)|", "", text) 
	text = re.sub('"', '', text)
	#just the words from the articles
	text = text.replace('\n', " ")
	sens = re.split("[.!?]", text.lower())
	sen_dict = {}
	for sen in sens:
		length = len(re.findall("\w+", sen))
		if length > 0:
			sen_dict[sen] = length
	return sen_dict

def get_sen_df(text):
	sd = get_sen_dict(text)
	s = pd.Series(sd, name = "length")
	s.index.name = "sentence"
	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence 
					 	  #and the setence length respectively
	return df

text = extract_texts(99)
ky_df = get_sen_df(text)
```

```{r analyze_entropies}
get_entropies <- function(sen_length, df) {

  tokens <- df %>%
    filter(length == sen_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, sentence) %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n()) %>%
    group_by(word_order, word) %>%
    filter(word_order <= length)

  tokens %>%
    summarise(n = n()) %>%
    tidyboot(summary_function = function(x) x %>% 
               summarise(entropy = entropy(n, unit = "log2")),
             statistics_functions = function(x) x %>%
             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
    mutate(length = sen_length)
}

entropies <- map(6:12, ~get_entropies(.x, py$ky_df)) %>%
  bind_rows()

ggplot(entropies, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)
```