### 2018-10-04 wiki_entropy.Rmd

---
title: Wikipedia Text Mining and Entropy Analysis
author: Josef Klafka
date:
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
---

```{r setup, include=FALSE}
# load packages
library(DBI)
library(knitr)
library(tidyverse)
library(directlabels)
library(dplyr)

library(SnowballC) #stemmer
library(tidytext)
library(entropy)
library(tidyboot)
library(tokenizers)
library(tau)
library(gtools)
library(reticulate)

library(feather)

knitr::opts_chunk$set(echo = TRUE)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 16))
use_python("~/../../Library/Frameworks/Python.framework/Versions/3.7/bin/python3.7")
```

```{python, include=FALSE, eval = T}
import regex as re 
import numpy as np
import pandas as pd 
import csv

def extract_texts(upper_bound):
	rs = ""
	for i in range(upper_bound):
		if i < 10:
			number = "0" + str(i)
		else:
			number = str(i)
		f = open("../misc/text/AA/wiki_" + number, 'r')
		text = f.read()
		rs += text
	return rs

def get_sen_dict(text):
	text = re.sub("<doc.+>|</doc>|http:\S+|[-%,;:–'&*#/—“»]|\d+|\(|\)|", "", text) 
	text = re.sub('"', '', text)
	#just the words from the articles
	text = text.replace('\n', " ")
	sens = re.split("[.!?]", text.lower())
	sen_dict = {}
	for sen in sens:
		length = len(re.findall("\w+", sen))
		if length > 0:
			sen_dict[sen] = length
	return sen_dict

def get_sen_df(text):
	sd = get_sen_dict(text)
	s = pd.Series(sd, name = "length")
	del sd
	s.index.name = "gloss"
	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence 
					 	  #and the setence length respectively
	df["speaker_role"] = "not child"
	return df

text = extract_texts(20)
xmf_df = get_sen_df(text)

xmf_df.to_csv("xmf_df.csv")
```

```{r slope_code, include=FALSE, eval = F}
# extract_stem_tokens <- function(role, min_length, utterances) {
#   
#   REPS <- 100
#   
#   if (role == "child") {
#     lengths <- utterances %>%
#       filter(speaker_role == "Target_Child") %>%
#       mutate(length = str_count(stem, " ") + 1)
# 
#   } else {
#     lengths <- utterances %>%
#       filter(speaker_role != "Target_Child") %>%
#       mutate(length = str_count(stem, " ") + 1)
#   }
#   
#   tokens <- lengths %>% 
#     filter(length >= min_length) %>%
#     mutate(utterance_id = 1:n()) %>%
#     unnest_tokens(word, stem) %>%
#     group_by(utterance_id) %>%
#     mutate(word_order = 1:n()) %>%
#     group_by(length, word_order, word) %>%
#     summarise(n = n()) 
# }

extract_wiki_tokens <- function(min_length, max_length, utterances) {

  tokens <- utterances %>% 
    filter(length >= min_length) %>%
    filter(length <= max_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, gloss) %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n()) %>%
    group_by(length, word_order, word) %>%
    filter(word_order <= length) %>% 
    summarise(n = n()) 
}

extract_tokens <- function(min_length, max_length, utterances) {

  tokens <- utterances %>% 
    mutate(lengths = str_count(gloss, " ")) %>%
    filter(lengths >= min_length) %>%
    filter(lengths <= max_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n()) %>%
    group_by(length, word_order, word) %>%
    filter(word_order <= length) %>% 
    summarise(n = n()) 
}

start_slope <- function(df) {

 df %>%
    filter(word_order %in% c(1,2)) %>%
    sample_frac(1, replace = T) %>%
    group_by(length, word_order) %>%
    summarise(entropy = entropy(n, unit = "log2")) %>%
    spread(word_order, entropy) %>%
    ungroup() %>%
    mutate(diff = `2` - `1`) %>%
    summarise(diff = mean(diff)) %>%
    pull(diff)

}

second_slope <- function(df) {

 df %>%
    ungroup() %>%
    mutate(word_order %in% c(2, 3)) %>%
    filter(word_order > 0) %>%
    sample_frac(1, replace = T) %>%
    group_by(length, word_order) %>%
    summarise(entropy = entropy(n, unit = "log2")) %>%
    spread(word_order, entropy) %>%
    ungroup() %>%
    mutate(diff = `3` - `2`) %>%
    summarise(diff = mean(diff)) %>%
    pull(diff)

}

#note this measurement only makes sense if utterance length is at least 6
mid_slope <- function(df) {

 df %>%
    ungroup() %>%
    mutate(word_order = if_else(word_order == length-2, 2,
                                if_else(word_order == 3, 1,
                                        0))) %>%
    filter(word_order > 0) %>%
    group_by(length, word_order) %>%
    sample_frac(1, replace = T) %>%
    group_by(length, word_order) %>%
    summarise(entropy = entropy(n, unit = "log2")) %>%
    spread(word_order, entropy) %>%
    ungroup() %>%
    mutate(diff = `2` - `1`) %>%
    summarise(diff = mean(diff)) %>%
    pull(diff)

}

penult_slope <- function(df) {

 df %>%
    ungroup() %>%
    mutate(word_order = if_else(word_order == length-1, 2,
                                if_else(word_order == length -2, 1,
                                        0))) %>%
    filter(word_order > 0) %>%
    group_by(length, word_order) %>%
    sample_frac(1, replace = T) %>%
    group_by(length, word_order) %>%
    summarise(entropy = entropy(n, unit = "log2")) %>%
    spread(word_order, entropy) %>%
    ungroup() %>%
    mutate(diff = `2` - `1`) %>%
    summarise(diff = mean(diff)) %>%
    pull(diff)

}

end_slope <- function(df) {

 df %>%
    ungroup() %>%
    mutate(word_order = if_else(word_order == length, 2,
                                if_else(word_order == length -1 , 1,
                                        0))) %>%
    filter(word_order > 0) %>%
    group_by(length, word_order) %>%
    sample_frac(1, replace = T) %>%
    group_by(length, word_order) %>%
    summarise(entropy = entropy(n, unit = "log2")) %>%
    spread(word_order, entropy) %>%
    ungroup() %>%
    mutate(diff = `2` - `1`) %>%
    summarise(diff = mean(diff)) %>%
    pull(diff)

}

get_wiki_slopes <- function(min_length, max_length, utterances) {

  REPS <- 100

  tokens <- extract_tokens(min_length, max_length, utterances)

  start_slopes <- replicate(REPS, start_slope(tokens)) %>% 
    as_data_frame() %>%
    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
              ci_lower = quantile(value, .025))

  second_slopes <- replicate(REPS, second_slope(tokens)) %>% 
    as_data_frame() %>%
    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
              ci_lower = quantile(value, .025))

  mid_slopes <- replicate(REPS, mid_slope(tokens)) %>% 
    as_data_frame() %>%
    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
              ci_lower = quantile(value, .025))

  penult_slopes <- replicate(REPS, penult_slope(tokens)) %>% 
    as_data_frame() %>%
    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
              ci_lower = quantile(value, .025))

  end_slopes <- replicate(REPS, end_slope(tokens)) %>% 
    as_data_frame() %>%
    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
              ci_lower = quantile(value, .025))

  rbind(start_slopes, second_slopes, mid_slopes, penult_slopes, end_slopes)
}

```

```{r slope_analysis, eval = F}

xmf_df <- read_csv("~/Documents/Labwork/Joe/Positional-Information/xmf_df.csv")

#get_slopes("not child", 6, py$xmf_df, "gloss")
gtx <- extract_tokens(6, 25, xmf_df) %>% filter(word != " ") %>% filter(word != "")

gtx %>%
    filter(word_order %in% c(1,2)) %>%
    group_by(length, word_order) %>%
    summarise(entropy = entropy(n, unit = "log2"), n = n()) %>%
    spread(word_order, entropy) %>%
    group_by(length) %>%
    mutate(diff = `2` - `1`) %>%
  ggplot(aes(x = length, y = diff)) +
  geom_point(aes(size = n)) +
  geom_smooth(se = F) +
  scale_x_log10() #log length stats

get_wiki_slopes(min_length = 6, max_length = 50, utterances=xmf_df)

second_slopes <- replicate(100, second_slope(gtx)) %>% 
    as_data_frame() %>% View()
    # summarise(mean = mean(value), ci_upper = quantile(value, .975), 
    #           ci_lower = quantile(value, .025))

print(second_slopes)
```

```{r analyze_entropies, include=FALSE, eval = F}
get_entropies <- function(sen_length, df) {

  tokens <- df %>%
    filter(length == sen_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, sentence) %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n()) %>%
    group_by(word_order, word) %>%
    filter(word_order <= length)

  tokens %>%
    summarise(n = n()) %>%
    tidyboot(summary_function = function(x) x %>% 
               summarise(entropy = entropy(n, unit = "log2")),
             statistics_functions = function(x) x %>%
             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
    mutate(length = sen_length)
}

#entropies <- map(6:12, ~get_entropies(.x, py$xmf_df)) %>%
#  bind_rows()

#ggplot(entropies, aes(x = word_order, y = empirical_entropy,
#                      ymin = ci_lower, ymax = ci_upper)) +
#  facet_wrap(~ length) + 
#  geom_pointrange() +
#  geom_smooth(se = F)
```