### 2018-10-05 slope_comparison.Rmd
new cleaner file to test slope numbers on

---
title: "Untitled"
author: "Joe Klafka"
date: "10/5/2018"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(directlabels)
library(childesr) #data
library(tidytext)
library(entropy)
library(tidyboot)
library(dplyr)
library(tokenizers)
library(gtools)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 16))
```

```{r data}
prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss

xmf_df <- read_csv("xmf_df.csv")

```

```{r slope_analysis}

extract_childes_tokens <- function(role, min_length, max_length, utterances) {

  if (role == "child") {
        lengths <- utterances %>%
            filter(speaker_role != "Target_Child") %>%
            mutate(length = str_count(gloss, " ") + 1)
  } else {
        lengths <- utterances %>%
            filter(speaker_role != "Target_Child") %>%
            mutate(length = str_count(gloss, " ") + 1)
  }

  sub_tokens <- lengths %>% 
     mutate(gloss = str_trim(gloss)) %>%
     mutate(length = str_count(gloss, " +") +1) %>%
     filter(length >= min_length) %>%
     filter(length <= max_length) %>%
     mutate(utterance_id = 1:n()) %>%
     unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
     group_by(utterance_id) %>%
     mutate(word_order = 1:n()) %>%
     group_by(length, word_order, word) %>%
     filter(word_order <= length) %>% 
     summarise(n = n()) 
}

extract_wiki_tokens <- function(min_length, max_length, utterances) {

  tokens <- utterances %>% 
    mutate(gloss = str_trim(gloss)) %>%
    mutate(length = str_count(gloss, pattern = " +") + 1) %>% 
    filter(length >= min_length) %>%
    filter(length <= max_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n()) %>%
    # group_by(utterance_id) %>%
    group_by(length, word_order, word) %>%
    # filter(word_order <= length) %>% 
    summarise(n = n()) # %>%
    # left_join(utterances) %>%
    # select(gloss, n, length) %>%
    # View()
}

length_comparison <- function(min_length, max_length, utterances) {

  sub_tokens <- utterances %>% 
     mutate(gloss = str_trim(gloss)) %>%
     mutate(length = str_count(gloss, " +") +1) %>%
     filter(length >= min_length) %>%
     filter(length <= max_length) %>%
     mutate(utterance_id = 1:n())

sub_tokens %>%
     unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
     group_by(utterance_id) %>%
     mutate(word_order = 1:n()) %>%
     group_by(utterance_id) %>%
     summarise(n = n()) %>%
     # filter(n != 6) %>%
     left_join(sub_tokens) %>%
     select(gloss, n, length) %>%
     View()
}

n_comparison <- function(tokens) {

  tokens %>%
    filter(word_order %in% c(1,2)) %>%
    dplyr::group_by(length, word_order) %>%
    dplyr::summarise(entropy = entropy(n, unit = "log2"), n = n()) 
}

```

```{r slope_work}
prov_tokens <- extract_childes_tokens("not child", 6, 30, prov_utterances)

xmf_tokens <- extract_wiki_tokens(6, 25, xmf_df) %>% filter(word != " ") %>% filter(word != "")

```

```{r leftover code}
tokens %>%
    filter(word_order %in% c(1,2)) %>%
    dplyr::group_by(length, word_order) %>%
    dplyr::summarise(entropy = entropy(n, unit = "log2"), n = n()) %>%
    spread(word_order, entropy) %>%
    group_by(length) %>%
    mutate(diff = `2` - `1`) %>%
  ggplot(aes(x = length, y = diff)) +
  geom_point(aes(size = n)) +
  geom_smooth(se = F) +
  scale_x_log10() #log length stats

```