### 2019-01-25 wiki_entropy/wiki_clustering.Rmd

---
title: Wikipedia Text Mining and Entropy Analysis
author: Josef Klafka and Dan Yurovsky
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
---

```{r setup, include=FALSE}
library(tidyverse)
library(googlesheets)
library(lingtypology)
library(knitr)
library(missMDA)
library(cluster)
library(widyr)
library(lsa)
library(magrittr)
library(broom)

knitr::opts_chunk$set(echo = TRUE)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 16))
```

```{r hierarchical clustering}
NUM_CLUSTERS <- 10

##read in googlesheet
wiki_gs <- gs_title("Wikipedia_absolute")
wkpd <- gs_read(wiki_gs) 
LANGUAGES <- wkpd$Language

wkpd <- wkpd %>% 
  as.data.frame() %>%
  column_to_rownames(var = "Language")

##do clustering
wk_clust_tree <- wkpd %>% 
              dist() %>%
              hclust("ward.D2")

wk_clust_tree %>% plot(hang = -1)
wk_clusters <- wk_clust_tree %>%
              cutree(k = NUM_CLUSTERS)

wk_clusters %>% 
  t() %>%
  as_data_frame() %>% 
  t() %>%
  as_data_frame(rownames = "language") %>%
  rename(cluster = V1) %>%
  arrange(cluster) %>% 
  View()
```

```{r}
wiki_gs <- gs_title("Wikipedia_absolute")
wkpd <- gs_read(wiki_gs) 
LANGUAGES <- wkpd$Language

FEATURES <- c(1:144)
for (i in c(1:144)) {
  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
} #get all WALS features in correct syntax

wf <- wals.feature(FEATURES)

wf_df <- wf %>%
  as_data_frame() %>%
  filter(language %in% LANGUAGES) %>%
  group_by(language) %>%
  slice(1) %>%
  gather(feature, value, `1A`:`144A`) %>%
  group_by(feature) %>%
  mutate(value = factor(value))

cosines <- wkpd %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, cosine, -language1)

```

```{r}

wf_many_langs <- wf_df %>%
  group_by(feature) %>%
  summarise(n = sum(!is.na(value))) %>%
  filter(n > 60)

wf_sub_langs <- wf_df %>%
  filter(feature %in% wf_many_langs$feature) %>%
  group_by(language) %>%
  summarise(keep = sum(is.na(value)) == 0) %>%
  filter(keep)

wf_sub <- wf_df %>%
  filter(feature %in% wf_many_langs$feature,
         language %in% wf_sub_langs$language) %>%
  select(language, feature, value) %>%
  mutate(value = as.numeric(as.factor(value)))

wf_pairwise <- expand.grid(language1 = wf_sub$language,
                           language2 = wf_sub$language) %>%
  left_join(wf_sub, by = c("language1" = "language")) %>%
  rename(value1 = value) %>%
  left_join(wf_sub, by = c("language2" = "language", "feature")) %>%
  rename(value2 = value) %>%
  mutate(same = value1 == value2) %>%
  select(-value1, -value2) %>%
  left_join(cosines, by = c("language1", "language2"))

models <- wf_pairwise %>%
  group_by(feature) %>%
  nest() %>%
  mutate(model = map(data, ~glm(same ~ cosine, 
                                family = "binomial", data = .)))

models %>%
  mutate(coeffs = map(model, tidy)) %>%
  select(-data, -model) %>%
  unnest() %>%
  filter(term == "cosine") %>%
  arrange(desc(statistic))
```

```{r imputation for glm}
imputed_result <- wf_df %>%
  select(language, feature, value) %>%
  ungroup() %>% 
  spread(feature, value) %>% 
  column_to_rownames("language") %>%
  sapply(as.factor) %>%
  as_data_frame() %>%
  MIMCA(ncp = 7, maxiter = 999)

imputed_wf <- as.data.frame(imputed_result[[1]][[99]])
```

```{r evaluate_clusters}
wals_clusters <- wals_clust["cluster"] %>% unlist(use.names= FALSE)
```