### 2019-02-01 cogsci/position-info.tex

% Template for Cogsci submission with R Markdown

% Stuff changed from original Markdown PLOS Template
\documentclass[10pt, letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}
\usepackage{caption}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}

% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{apacite}

% KM added 1/4/18 to allow control of blind submission

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% % Text layout
% \topmargin 0.0cm
% \oddsidemargin 0.5cm
% \evensidemargin 0.5cm
% \textwidth 16cm
% \textheight 21cm

\title{Information Distribution Depends on Language-Specific Features}

\author{Josef Klafka \and Daniel Yurovsky \\
        \texttt{\{jklafka, yurovsky\}@uchicago.edu} \\
       Department of Psychology \\ University of Chicago}

\begin{document}

\maketitle

\begin{abstract}
Although languages vary widely in their structure, all are vehicles for
the transmission of information. Consequently, aspects of speakers' word
choice can be understood through the lens models of optimal
communication. One prediction of these models is that speakers should
keep the density of information constant over the duration of an
utterance (Levy \& Jaeger, 2007). However, different languages have
different coding models that constraint the space of speaker's choices
(e.g.~canonical word order). We build on a mdethod developed by Yu,
Cong, Liang, \& Liu (2016) to analye the entropy curves of natural
language productions across a diverse set of languages and in diverse
written and spoken contexts. We show languages have characteristic
constraints that they impose on speaker's choices that predict
deviations from Uniform Information Density, and that cross-linguistic
variability in these deviations is predictable in part from syntactic
properties of those languages.

\textbf{Keywords:}
Uniform information density; language structure; corpus analysis
\end{abstract}

Over 7,000 languages are spoken around the modern world (Simons \&
Fenning, 2018). These language vary along many dimensions, but all share
a core goal: communicating information. If speakers and writers of these
languages act near-optimally to achieve their communicative goals,
regularities of use across these diverse languages can be explained by a
rational theory of communication (Anderson, 1991). Information theory, a
mathematical framework developed by Shannon (1948) to describe the
transmission and decoding of signals, has been a unifying language for
the recent development of such theories in human and machine language
processing (Jelinek, 1976; Levy \& Jaeger, 2007).

These theories model the process of communication as transmission of
information over a noisy channel. The producer begins with an intended
meaning, packages this meaning into language, and then sends the meaning
to their intended receiver over a communcative channel. The receiver
must then decode from the signal they receive on their end of the
channel the producer's intended meaning. The problem is that the channel
is noisy, and sometimes the signal can get corrupted (e.g.~the producer
can misspeak, or the receiver can mishear). In order to maximize the
probability that the correct meaning is transmitted, these theories
predict that producers should choose linguistic messages that keep the
rate of information across words constant. The intuition is that if the
receiver misperceives a word, and that word contains most of the
information in the sentence, then the communication will have failed.
Because producers cannot predict which word a speaker will mishear,
their best strategy is spread the information evenly across all of the
words in a sentence, i.e.~maintain \emph{uniform information density}
(Genzel \& Charniak, 2002; Levy \& Jaeger, 2007).

The original evidence in Levy \& Jaeger (2007) finds that the insertion
of complementizers (e.g. ``that'') in relative clauses in English
corresponds to where neighboring words have high information content.
Similarly, Frank \& Jaeger (2008) argues that contradictions in English
such as ``you're'' do not occur when neighboring words are highly
informative. The evidence in favor of Uniform Information Density
largely been situation-specific and English-language driven, while the
hypothesis itself has been applied broadly over the past decade.
Applications include determining whether linguistic alignment takes
place (Jaeger \& Snider, 2013), Zipfian word length distributions
(Piantadosi, Tily, \& Gibson, 2011), communication efficiency (Mahowald,
Fedorenko, Piantadosi, \& Gibson, 2013), dialogue and turn-taking (Xu \&
Reitter, 2018) and the significance of ambiguity in language
(Piantadosi, Tily, \& Gibson, 2012), among other research.

However, other recent work has contradicted the Uniform Information
Density hypothesis. Similar to the original Levy \& Jaeger (2007), Zhan
\& Levy (2018) focuses on information distribution at particular points
in sentences. Zhan \& Levy (2018) finds that more information-rich
classifiers in Mandarin Chinese are produced when production of the
neighboring noun is difficult, not when the information content is high.
Jain, Singh, Ranjan, Rajkumar, \& Agarwal (2018) examine word order
across spoken sentences in Hindi, a freer word order language than
English, and find that information density has no significant effect on
word order.

Recently, Yu et al. (2016) developed a more direct test of the Uniform
Information Density hypothesis, applying the logic used by Genzel \&
Charniak (2002) to look at the distribution of information \emph{within}
individual sentences. Because people process language
incrementally--using the previous words in a sentence to predict the
words that will come next--the amount of information that a word
contains when seen in isolation should increase over the course of a
sentence (Ferrer-i-Cancho, Debowski, \& Prado Martin, 2013). Analyzing a
large corpus of written English, they find a different pattern: Entropy
increases over the first few words of an utterance and then remains
constant until the final word where it again jumps up (see top of Figure
\ref{fig:read_and_plot_exp1}). Yu et al. (2016) conclude that the
Uniform Information Density hypothesis must not hold for medial words in
a sentence.

We extend and generalize Yu et al. (2016) in three ways: We confirm that
this same pattern is found in spoken English--both in written language,
and in conversational speech between adults and between parents and
their children. Thus, this entropy curve is a robust feature of English
productions. We then examine entropy curves cross-linguistically, and
show that characteristic curves vary across langauge the world's
languages. Finally, we show that this variation is predictable in part
from the structure of individual languages (i.e.~word order). Taken
together, our results suggest a refinement of the Uniform Information
Density hypothesis: speakers may structure their utterances to optimize
information density, but they must do so under the predictable
constraints of their language.

\hypertarget{calculating-entropy-curves}{%
\subsection{Calculating entropy
curves}\label{calculating-entropy-curves}}

In all oll of our studies, we used an adaptation of the by-word entropy
method developed by Yu et al. (2016). Given a text or speech corpus
divided into individual utterances, we partition the corpus by utterance
length in number of words. For each word position \(X\) of utterances of
length \(k\), we define \(w\) as a unique word occurring in position
\(X\). We further define \(p(w)\) as the number of times word \(w\)
occurs in position \(X\), divided by the number of total words that
occur in position \(X\) i.e.~the number of sentences of length \(k\).
This creates a probability distribution over the words occurring in
position \(X\), and computing the Shannon (1948) entropy \(H(X)\) of
this probability distribution gives the positional entropy of position
\(X\) in utterances of length \(k\).

\[H(X) = \sum\limits_w p(w)\log\big(p(w)\big)\]

With this measure, we compute the unigram entropy at each position of
sentences of each length within the corpus. The result of this method
can be plotted for each utterance length as an \emph{entropy curve},
which can be visually compared across utterance length to observe the
how the unigram entropy changes across absolute positions in each of the
utterances. Genzel \& Charniak (2002) similarly examine a unigram
entropy measure on sentences, and found that entropy at the sentence
level increases linearly with sentence index within a corpus. Uniform
Information Density applies this uniformity of entropy rate in sentences
to all levels of speech, and so our method obtained from Yu et al.
(2016), which examines text at the word level, should find a
monotonically increasing at the word level.

The entropy curves capture individual variation across positions in
utterances of the same length. This allows us to directly observe and
judge the amount of variation in words that appear in an individual
position of a sentence. We can directly compare any two positions within
utterances to determine the amount of uncertainty, and therefore
information, on average contained by words within that position of
utterances. This method is thus identical in logic to Genzel \& Charniak
(2002), but within sentences instead of across sentences.

\hypertarget{study-1}{%
\section{Study 1}\label{study-1}}

We began by replicating Yu et al.'s (2016) analysis, computing entropy
curves on the British National Corpus--a collection of predominantly
written English (Clear, 1993). We then applied the same method to the
Switchboard corpus--a collection of spoken language (Godfrey, Holliman,
\& McDaniel, 1992). This allows us to ask whether the characteristic
function identified by Yu et al. (2016) is a general feature of English,
or instead a function of written language.

\hypertarget{data-and-analysis}{%
\section{Data and Analysis}\label{data-and-analysis}}

The British National Corpus consists of predominantly (90\%) written
language documents collected in the early 1980s and 1990s across a
variety of genres (scientific articles, newspapers, fiction, etc). It
also contains a small collection of spoken language. All together, it
contains \(\sim\) 100 million tokens. The Switchboard corpus is a
collection of \(\sim\) 2,400 telephone conversations between
unacquainted adults prompted with subjects of conversation. Switchboard
is the corpus used in Levy and Jaeger's (2007) original demonstration of
the Uniform Information Density Hypothesis.

For each corpus, we computed entropy curves using the method described
above for all sentences from length 4 to 30.

\hypertarget{results}{%
\section{Results}\label{results}}

The entropy curves computed in both the British National Corpus and
Switchboard were remarkably consistent both across corpora and across
the range of sentence lenghts we analyzed (Figure
\ref{fig:read_and_plot_exp1}). Confirming Yu et al.'s (2016) findings,
we find that The positional entropy at the beginning of sentences is
low, then rises and plateaus for sentence-medial word positions before
dropping slightly in the second-to-last position and rising again.

\begin{CodeChunk}
\begin{figure*}[tb]

{\centering \includegraphics{figs/read_and_plot_exp1-1}

}

\caption[Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom)]{Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom). Points average entropies, error bars show 95\% confidence intervals computed by non-parametric bootstrap.}\label{fig:read_and_plot_exp1}
\end{figure*}
\end{CodeChunk}

The shape of positional entropy curves we find in these two corpora is
notable for two reasons: (1) it does not follow our predictions from
Uniform Information Density, and (2) the distribution is robust across
written and spoken Engilsh. This suggests that the entropy curve is
characteristic of the English language. We next asked whether this shape
is a feature even of children's speech, and also whether it varies
cross-linguistically.

\hypertarget{study-2}{%
\section{Study 2}\label{study-2}}

To understand how robust this entropy curve is, we turned to
conversational speech from parent-child interactions. If we find the
same shape even in children's productions, we have even stronger
evidence that the 3-step curve is a charcateristic feature of English.
We thus turned to the Child Language Data Exchange System (CHILDES), a
collection of transcripts of parent-child interactions (MacWhinney,
2014).

\hypertarget{data-and-analysis-1}{%
\subsection{Data and Analysis}\label{data-and-analysis-1}}

We analyzed three seperate corpora in CHILDES: The Providence Corpus
(Demuth, Culbertson, \& Alter, 2006), The Shiro corpus (Shiro, 2000),
and the Zhou Dinner Corpus (Li \& Zhou, 2015). The Providence corpus
consists of conversations between six 1--3-year-old American English
speaking children and their parents recorded in their homes. Providence
corpus recorded interactions between children between 1 and 3 years old
and their parents in the home. The Shiro Corpus consists of prompted
Spanish-language narratives individually collected from over a hundred
Venezualan schoolchildren, half from high SES backgrounds and half from
low SES backgrounds. The Zhou Dinner Corpus contains dinner
conversations between 5 to 6-year-old Mandarin speaking children and
their parents collected in Shanghai. Spanish is an Indo-European
language like English, possessing similar grammar, word order and
numerous cognate word. In contrast, Mandarin Chinese is typologically
completed unrelated to English.

We accessed the transcripts from each corpus using \texttt{childesr}, an
R-interface to a database formatted version of CHILDES (Sanchez et al.,
in press). We divided all utterances from each transcript into those
produced by target child, and those produced by all other speakers. We
then applied the same entropy curve method described above. For
Mandarin, we used pinyin transliterations of the utterances in the
corpus with demarcated word boundaries. The Chinese characters used for
writing Mandarin do not normally demarcate word boundaries by spacing
words apart, and for normal Chinese writing including spaces between
word boundaries can have a negative effect on reading times (Bai, Yan,
Liversedge, Zang, \& Rayner, 2008).

\hypertarget{results-and-analysis}{%
\subsection{Results and Analysis}\label{results-and-analysis}}

\begin{CodeChunk}
\begin{figure*}[tb]

{\centering \includegraphics{figs/plot_childes-1}

}

\caption[Representative Entropy Curves for Three Childes Corpora in English, Mandarin, and Spanish]{Representative Entropy Curves for Three Childes Corpora in English, Mandarin, and Spanish. Points average entropies, error bars show 95\% confidence intervals computed by non-parametric bootstrap.}\label{fig:plot_childes}
\end{figure*}
\end{CodeChunk}

Across corpora, we found similar entropy curve shapes for adults and
children, but distinct shapes for each language. Figure
\ref{fig:plot_childes} shows representative curves across corpora, but
shapes were robust across the full range utterances we analyzed. We
found a distinct three-step distribution for English and Spanish CHILDES
corpora, with a slight dip in the penultimate position of each sentence.
The Mandarin corpus entropy curve, by comparison, has a noticeably lower
positional entropy values in utterance-final positions than in
utterance-penultimate positions.

These results present two important pieces of information. First, our
analysis of the Providence corpus broadly replicates the shape we found
in both the British National Corpus and in Switchboard. Thus, the
entropy curve of English appears not just in adult-adult conversation,
but even in speech produced by parents to their children, and speech
produced by very young children to their parents. This suggests that it
is a highly robust feature of English langauge, as it structures the
productions of even pre-schooled aged children.

Second, the Entropy curves of English, Spanish, and Mandarin were not
identical. None of these shapes resembled the monotonically increasing
function predicted naively from Uniform Information Density, but also
Mandarin was quite different from both English and Spanish. This
suggests that the entropy curve can vary from language to language. One
possibility is that this variation arises from typological features of
languages, such as characteristic word order. We noticed, for instance,
a relatively high density of determiners in the penultimate position of
English and Spanish utterances, which could account for the penultimate
dip in those languages. In our final analysis, we explored this
possibility directly, analyzing a large set of Wikipedia corpora from
diverse languages, and asking whether variability in their entropy
curves was related to variability in their syntactic structure.

\hypertarget{experiment-3}{%
\section{Experiment 3}\label{experiment-3}}

The UID hypothesis also applies to written communication: we expect
people to communicate information at a uniform rate through writing as
well. We use Wikipedia as a source for written data, which provides two
advantages. One, the quantity of data in Wikipedia is very large for
each language, at least several megabytes, and two, there are hundreds
of languages with their own Wikipedia corpora. We can therefore run our
entropy analysis on a much greater scale than using small spoken CHILDES
corpora and directly compare the results of the entropy analysis on each
language to one another. We expect to find linguistic features that
determine the shape of a language's entropy curve. We treated sentences
in the Wikipedia text corpora as equivalent to utterances in the
Switchboard and CHILDES corpora.

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

Using Giuseppe Attardi's Wikiextractor tool
\footnote{https://github.com/attardi/wikiextractor}, we extract text
corpora from Wikipedia by downloading a stored collection of Wikipedia
entries in each langauge and randomly selecting several thousand
articles from each Wikipedia language. Each language corpus was cleaned
and limited to sentences between six and \(50\) words. Similar to our
process for the Switchboard and CHILDES spoken corpora, we divided each
corpus by sentence length, and then computed our positional entropy
measure on each word position within each sentence length.

We computed two slope treatments of each curve. In the \emph{absolute}
treatment, with sentence length denoted as \(k\), we computed the slope
between positions \(1\) and \(2\), positions \(2\) and \(3\), positions
\(3\) and \(k-2\), positions \(k-2\) and \(k-1\) and positions \(k-1\)
and \(k\). For the short utterances appearing the CHILDES speech
corpora, the slopes between the first and second slopes on either end of
the distributions appeared to be more characteristic of the distribution
as a whole, with a plateau in the middle of the entropy curve for each
of the language corpora we examined in CHILDES.

However, sentences of length greater than ten in the Wikipedia corpora
were significantly more common than in the CHILDES. Therefore we also
computed relative slope treatments. In the \emph{relative 5} treatment,
we computed the slopes between every \(20\%\) of the relative word
positions in each sentence length, e.g.~the slope between \(0\%\) and
\(20\%\) of sentences of length \(10\) is the slope between the position
entropies at the sentence-initial word position and the position
representing the third word. When computing the word position index to
make those cuts, if the slope was not a whole number, then the closest
whole number position was used instead for slope calculation. Each
comparative slope within each treatment was averaged together between
different sentence lengths, for example all of the \(0\%\) to \(20\%\)
slopes over all sentence lengths were averaged together.

In both the absolute and relative \(5\) treatments, each language is
embedded in \(5\)-dimensional space. This permitted us to use cosine
similarity for direct comparison between the languages we pulled from
Wikipedia. To determine which phonological, morphological and syntactic
features affected the embedding of a language in the Wikipedia dataset,
we use the linguistic features in World Atlas of Language Structures
Dryer \& Haspelmath (2013). We checked the effects of individual
features on the embeddings of languages in the different treatments. We
computed pairwise cosine similarity between each pair of language
vectors within each treatment. For a subset of eight WALS features, we
used a generalized linear model to see whether the cosine similarity
between languages mattered in predicting if the languages shared the
same value for a WALS feature.

\hypertarget{results-and-analysis-1}{%
\subsection{Results and Analysis}\label{results-and-analysis-1}}

For eight features and \(45\) languages we pulled from Wikipedia, we
found that the cosine similarity between WALS features. The table below
shows the results for the generalized linear model we computed using
WALS features and the absolute treatment.

\begin{table}[tb]
\centering
\begin{tabular}{llrrrr}
  \hline
feature & term & estimate & std.error & statistic & p.value \\ 
  \hline
83A & cosine & 1.84 & 0.01 & 178.22 & 0.00 \\ 
  95A & cosine & 1.90 & 0.01 & 170.98 & 0.00 \\ 
  81A & cosine & 1.51 & 0.01 & 153.70 & 0.00 \\ 
  97A & cosine & 1.58 & 0.01 & 136.81 & 0.00 \\ 
  144A & cosine & 0.88 & 0.01 & 74.43 & 0.00 \\ 
  138A & cosine & 0.40 & 0.01 & 46.12 & 0.00 \\ 
  87A & cosine & 0.33 & 0.01 & 39.91 & 0.00 \\ 
  143A & cosine & 0.37 & 0.01 & 39.74 & 0.00 \\ 
  82A & cosine & 0.03 & 0.01 & 3.23 & 0.00 \\ 
   \hline
\end{tabular}
\end{table}

We also used the linear model approach to evaluate the effects of cosine
distance in the relative \(5\) treatment on the values of the same WALS
features and found similar outcomes.

From these results, we argue that cosine distance played some role in
the feature determination. The top four features in determining the WALS
feature value all characterize word order, which indicates that for this
subset of features and languages that the embeddings in the slope space
are related to the WALS features. We argue therefore that typological
features play a role in determining the positional entropy values for a
language. This indicates that the entropy curves are in some way
structured by the syntactic, morphological and phonological features of
a language.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

In this paper, we have extended and applied a model from Yu et al.
(2016) and derived from Genzel \& Charniak (2002) for distinguishing
entropy at each word position within sentences. We have argued that the
outcomes of this model are derived from the distribution of information
transmission in human communication. Language is not a uniform noisy
channel of communication, but each language individually represents a
different noisy channel of communication characterized in part by
typological features.

Our work complements the approach of studies such as Aylett \& Turk
(2004), where languages are characterized by a single number,
representing the rate of semantic information transfer per syllable.
This body of work attests to languages possessing different rates of
information transfer based on typological features as well. Within the
information transfer rate literature, the rate is an overall feature of
language that does not vary corpus to corpus. Similarly, in our work we
have obtained a characteristic entropy distribution for each language.
Both information rate and entropy distribution are characterized by
typological features. This indicates that the structure and rate of how
people communicate information varies cross-linguistically based on
typological features.

We initially wanted to use several hundred language corpora pulled from
Wikipedia and all \(144\) linguistic features from WALS. We considered
an computing unsupervized clusterings of the \(5\)-dimensional vectors
of our Wikipedia entropy curves, and comparing those clusterings to
clusterings of the WALS features directly using a cluster similarity
measure such as the Rand index Rand (1971). However, the problem then
arises of which combination of WALS features and how many features to
include in the clustering analysis, with \(144!\) different
combinations. Two additional problems presented themselves: one,
deciding how many clusters to use in an unsupervised clustering analysis
is an unsolved problem in machine learning; and two, not all of the
languages on Wikipedia have values inputted for their WALS features. As
a follow-up, we are considering missing-data imputation on the WALS
features in order to run the generalized linear model analysis on a
large dataset of languages.

On a more mechanical level, we expect that our entropy model will have
implications for how long people take to read individual words at
different positions of a sentence. Eye-tracking research has indicated
the effects of surprisal on fixation duration during eye-tracking
studies (Boston, Hale, Kliegl, Patil, \& Vasishth, 2008; Demberg \&
Keller, 2008; Smith \& Levy, 2013); higher surprisal of a word is on
average predictive of higher fixation duration. The \emph{wrap-up
effect} states that people process sentence-final words more slowly on
average then sentence-medial or sentence-initial words when reading, due
to integrating information to form a final understanding of the
sentence's meaning (Kuperman, Dambacher, Nuthmann, \& Kliegl, 2010;
Stowe, Kaan, Sabourin, \& Taylor, 2018). The wrap-up effect is drawn
from evidence in languages with a large final increase in their entropy
curve from our study, so we hypothesize that the supposed wrap-up effect
derives from the same source sentence-final increase in entropy curves
observed in English and other Germanic languages.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

{[}Not here yet.{]}

\hypertarget{references}{%
\section{References}\label{references}}

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}

\noindent

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-anderson1991}{}%
Anderson, J. R. (1991). The adaptive nature of human categorization.
\emph{Psychological Review}, \emph{98}(3), 409.

\leavevmode\hypertarget{ref-aylett2004}{}%
Aylett, M., \& Turk, A. (2004). The smooth signal redundancy hypothesis:
A functional explanation for relationships between redundancy, prosodic
prominence, and duration in spontaneous speech. \emph{Language and
Speech}, \emph{47}(1), 31--56.

\leavevmode\hypertarget{ref-bai2008}{}%
Bai, X., Yan, G., Liversedge, S. P., Zang, C., \& Rayner, K. (2008).
Reading spaced and unspaced chinese text: Evidence from eye movements.
\emph{Journal of Experimental Psychology: Human Perception and
Performance}, \emph{34}(5), 1277.

\leavevmode\hypertarget{ref-boston2008}{}%
Boston, M. F., Hale, J., Kliegl, R., Patil, U., \& Vasishth, S. (2008).
Parsing costs as predictors of reading difficulty: An evaluation using
the potsdam sentence corpus. \emph{Journal of Eye Movement Research},
\emph{2}(1).

\leavevmode\hypertarget{ref-clear1993}{}%
Clear, J. H. (1993). The british national corpus. \emph{The Digital
World}, 163--187.

\leavevmode\hypertarget{ref-demberg2008}{}%
Demberg, V., \& Keller, F. (2008). Data from eye-tracking corpora as
evidence for theories of syntactic processing complexity.
\emph{Cognition}, \emph{109}(2), 193--210.

\leavevmode\hypertarget{ref-demuth2006}{}%
Demuth, K., Culbertson, J., \& Alter, J. (2006). Word-minimality,
epenthesis and coda licensing in the early acquisition of english.
\emph{Language and Speech}, \emph{49}(2), 137--173.

\leavevmode\hypertarget{ref-dryer2013}{}%
Dryer, M. S., \& Haspelmath, M. (2013). \emph{The world atlas of
language structures online}. Max Planck Institute for Evolutionary
Anthropology.

\leavevmode\hypertarget{ref-ferrer-i-cancho2013}{}%
Ferrer-i-Cancho, R., Debowski, L., \& Prado Martin, F. M. del. (2013).
Constant conditional entropy and related hypotheses. \emph{Journal of
Statistical Mechanics: Theory and Experiment}, \emph{2013}(07), L07001.

\leavevmode\hypertarget{ref-frank2008}{}%
Frank, A. F., \& Jaeger, T. F. (2008). Speaking rationally: Uniform
information density as an optimal strategy for language production. In
\emph{Proceedings of the annual meeting of the cognitive science
society} (Vol. 30).

\leavevmode\hypertarget{ref-genzel2002}{}%
Genzel, D., \& Charniak, E. (2002). Entropy rate constancy in text. In
\emph{Proceedings of the 40th annual meeting on association for
computational linguistics} (pp. 199--206). Association for Computational
Linguistics.

\leavevmode\hypertarget{ref-godfrey1992}{}%
Godfrey, J. J., Holliman, E. C., \& McDaniel, J. (1992). SWITCHBOARD:
Telephone speech corpus for research and development. In
\emph{Acoustics, speech, and signal processing, 1992. ICASSP-92., 1992
ieee international conference on} (Vol. 1, pp. 517--520). IEEE.

\leavevmode\hypertarget{ref-jaeger2013}{}%
Jaeger, T. F., \& Snider, N. E. (2013). Alignment as a consequence of
expectation adaptation: Syntactic priming is affected by the prime's
prediction error given both prior and recent experience.
\emph{Cognition}, \emph{127}(1), 57--83.

\leavevmode\hypertarget{ref-jain2018}{}%
Jain, A., Singh, V., Ranjan, S., Rajkumar, R., \& Agarwal, S. (2018).
Uniform information density effects on syntactic choice in hindi. In
\emph{Proceedings of the workshop on linguistic complexity and natural
language processing} (pp. 38--48).

\leavevmode\hypertarget{ref-jelinek1976}{}%
Jelinek, F. (1976). Continuous speech recognition by statistical
methods. \emph{Proceedings of the IEEE}, \emph{64}(4), 532--556.

\leavevmode\hypertarget{ref-kuperman2010}{}%
Kuperman, V., Dambacher, M., Nuthmann, A., \& Kliegl, R. (2010). The
effect of word position on eye-movements in sentence and paragraph
reading. \emph{The Quarterly Journal of Experimental Psychology},
\emph{63}(9), 1838--1857.

\leavevmode\hypertarget{ref-levy2007}{}%
Levy, R. P., \& Jaeger, T. F. (2007). Speakers optimize information
density through syntactic reduction. In \emph{Advances in neural
information processing systems} (pp. 849--856).

\leavevmode\hypertarget{ref-li2015}{}%
Li, H., \& Zhou, J. (2015). \emph{Study on dinner table talk of
preschool children family in shanghai} (Master's thesis). East China
Normal University.

\leavevmode\hypertarget{ref-macwhinney2014}{}%
MacWhinney, B. (2014). \emph{The childes project: Tools for analyzing
talk, volume ii: The database}. Psychology Press.

\leavevmode\hypertarget{ref-mahowald2013}{}%
Mahowald, K., Fedorenko, E., Piantadosi, S. T., \& Gibson, E. (2013).
Info/information theory: Speakers choose shorter words in predictive
contexts. \emph{Cognition}, \emph{126}(2), 313--318.

\leavevmode\hypertarget{ref-piantadosi2011}{}%
Piantadosi, S. T., Tily, H., \& Gibson, E. (2011). Word lengths are
optimized for efficient communication. \emph{Proceedings of the National
Academy of Sciences}, \emph{108}(9), 3526--3529.

\leavevmode\hypertarget{ref-piantadosi2012}{}%
Piantadosi, S. T., Tily, H., \& Gibson, E. (2012). The communicative
function of ambiguity in language. \emph{Cognition}, \emph{122}(3),
280--291.

\leavevmode\hypertarget{ref-rand1971}{}%
Rand, W. M. (1971). Objective criteria for the evaluation of clustering
methods. \emph{Journal of the American Statistical Association},
\emph{66}(336), 846--850.

\leavevmode\hypertarget{ref-sanchez2019}{}%
Sanchez, A., Meylan, S. C., Braginsky, M., MacDonald, K. E., Yurovsky,
D., \& Frank, M. (in press). Childes-db: A flexible and reproducible
interface to the child language data exchange system. \emph{Behavior
Research Methods}.

\leavevmode\hypertarget{ref-shannon1948}{}%
Shannon, C. E. (1948). A mathematical theory of communication.
\emph{Bell System Technical Journal}, \emph{27}(3), 379--423.

\leavevmode\hypertarget{ref-shiro2000}{}%
Shiro, M. (2000). Diferencias sociales en la construcción del" yo" y
del" otro": Expresiones evaluativas en la narrativa de niños caraqueños
en edad escolar. In \emph{Lengua, discurso, texto: I simposio
internacional de análisis del discurso} (pp. 1303--1318). Visor.

\leavevmode\hypertarget{ref-simons2018}{}%
Simons, G. F., \& Fenning, C. D. (Eds.). (2018). \emph{Ethnologue:
Languages of the world, 21st edition}. Dallas, Texas: SIL International.

\leavevmode\hypertarget{ref-smith2013}{}%
Smith, N. J., \& Levy, R. (2013). The effect of word predictability on
reading time is logarithmic. \emph{Cognition}, \emph{128}(3), 302--319.

\leavevmode\hypertarget{ref-stowe2018}{}%
Stowe, L. A., Kaan, E., Sabourin, L., \& Taylor, R. C. (2018). The
sentence wrap-up dogma. \emph{Cognition}, \emph{176}, 232--247.

\leavevmode\hypertarget{ref-xu2018}{}%
Xu, Y., \& Reitter, D. (2018). Information density converges in
dialogue: Towards an information-theoretic model. \emph{Cognition},
\emph{170}, 147--163.

\leavevmode\hypertarget{ref-yu2016}{}%
Yu, S., Cong, J., Liang, J., \& Liu, H. (2016). The distribution of
information content in english sentences. \emph{arXiv Preprint
arXiv:1609.07681}.

\leavevmode\hypertarget{ref-zhan2018}{}%
Zhan, M., \& Levy, R. (2018). Comparing theories of speaker choice using
a model of classifier production in mandarin chinese. In
\emph{Proceedings of the 2018 conference of the north american chapter
of the association for computational linguistics: Human language
technologies, volume 1 (long papers)} (Vol. 1, pp. 1997--2005).

\bibliographystyle{apacite}

\end{document}