### 2019-02-02 cogsci/position-info.tex

% Template for Cogsci submission with R Markdown

% Stuff changed from original Markdown PLOS Template
\documentclass[10pt, letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}
\usepackage{caption}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}

% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{apacite}

% KM added 1/4/18 to allow control of blind submission

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% % Text layout
% \topmargin 0.0cm
% \oddsidemargin 0.5cm
% \evensidemargin 0.5cm
% \textwidth 16cm
% \textheight 21cm

\title{Information Distribution Depends on Language-Specific Features}

\author{Josef Klafka \and Daniel Yurovsky \\
        \texttt{\{jklafka, yurovsky\}@uchicago.edu} \\
       Department of Psychology \\ University of Chicago}

\begin{document}

\maketitle

\begin{abstract}
Although languages vary widely in their structure, all are vehicles for
the transmission of information. Consequently, aspects of speakers' word
choice can be understood through the models of optimal communication.
One prediction of these models is that speakers should keep the density
of information constant over the duration of an utterance (Levy \&
Jaeger, 2007). However, different languages have different coding models
that constrain the space of a speaker's choices (e.g.~canonical word
order). We build on a method developed by Yu, Cong, Liang, \& Liu (2016)
to analyze the word-level entropy curves of natural language productions
across a diverse set of languages and in diverse written and spoken
contexts. We show languages impose characteristic constraints on
speaker's choices that predict deviations from Uniform Information
Density, and that cross-linguistic variability in these deviations is
predictable in part from syntactic properties of those languages.

\textbf{Keywords:}
Uniform information density; language structure; corpus analysis
\end{abstract}

Over 7,000 languages are spoken around the modern world (Simons \&
Fenning, 2018). These language vary along many dimensions, but all share
a core goal: communicating information. If speakers and writers of these
languages act near-optimally to achieve their communicative goals,
regularities of use across these diverse languages can be explained by a
rational theory of communication (Anderson, 1991). Information theory, a
mathematical framework developed by Shannon (1948) to describe the
transmission and decoding of signals, has been a unifying language for
the recent development of such theories in human and machine language
processing (Jelinek, 1976; Levy \& Jaeger, 2007).

These theories model the process of communication as transmission of
information over a noisy channel. The producer begins with an intended
meaning, packages this meaning into language, and then sends the encoded
meaning to their intended receiver over a communicative channel. The
receiver must then decode the producer's intended meaning from the
signal they receive on their end of the channel. The problem is that the
channel is noisy, and sometimes the signal can get corrupted (e.g.~the
producer can misspeak, or the receiver can mishear). In order to
maximize the probability that the correct meaning is transmitted, these
theories predict that producers should choose linguistic messages that
keep the rate of information constant across words. The intuition is
that if the receiver misperceives a word, and that word contains most of
the information in the sentence, then the communication will have
failed. Because producers cannot predict which word a speaker will
mishear, their best strategy is spread the information evenly across all
of the words in a sentence, i.e.~maintain \emph{Uniform Information
Density} (Genzel \& Charniak, 2002; Levy \& Jaeger, 2007).

The evidence in favor of Uniform Information Density largely been
situation-specific and English-language driven, while the hypothesis
itself has been applied broadly over the past decade. The original
evidence in favor of Uniform Information Density in Levy \& Jaeger
(2007) finds that the insertion of complementizers (e.g. ``that'') in
relative clauses in English corresponds to where neighboring words have
high information content. Similarly, Frank \& Jaeger (2008) argues that
contractions in English such as ``you're'' do not occur when neighboring
words are highly informative. Applications of Uniform Information
Density include determining whether linguistic alignment takes place
(Jaeger \& Snider, 2013), Zipfian word length distributions (Piantadosi,
Tily, \& Gibson, 2011), communication efficiency (Mahowald, Fedorenko,
Piantadosi, \& Gibson, 2013), dialogue and turn-taking (Xu \& Reitter,
2018) and the significance of ambiguity in language (Piantadosi, Tily,
\& Gibson, 2012), among other research.

However, other recent work has contradicted the Uniform Information
Density hypothesis. Similar to the original Levy \& Jaeger (2007), Zhan
\& Levy (2018) focuses on information distribution at particular points
in sentences. Zhan \& Levy (2018) finds that more information-rich
classifiers in Mandarin Chinese are produced when production of the
neighboring noun is difficult, not when the information content is high.
Jain, Singh, Ranjan, Rajkumar, \& Agarwal (2018) examine word order
across spoken sentences in Hindi, a freer word order language than
English, and find that information density has no significant effect on
determining a Hindi's speaker's word order.

Recently, Yu et al. (2016) developed a more direct test of the Uniform
Information Density hypothesis, applying the logic used by Genzel \&
Charniak (2002) to look at the distribution of information \emph{within}
individual sentences. Because people process language
incrementally--using the previous words in a sentence to predict the
words that will come next--the amount of information that a word
contains when seen in isolation should increase over the course of a
sentence (Ferrer-i-Cancho, Debowski, \& Prado Martin, 2013). Analyzing a
large corpus of written English, they find a different pattern: Entropy
increases over the first few words of an utterance and then remains
constant until the final word where it again jumps up (see top of Figure
\ref{fig:read_and_plot_exp1}). Yu et al. (2016) conclude that the
Uniform Information Density hypothesis must not hold for medial words in
a sentence.

We extend and generalize Yu et al. (2016) in three ways: We confirm that
this same pattern is found in spoken English--in written language, and
in conversational speech between adults and between parents and their
children. Thus, this entropy curve is a robust feature of English
productions. We then examine entropy curves cross-linguistically, and
show that characteristic curves vary across the world's languages.
Finally, we show that this variation is predictable in part from the
structure of individual languages (i.e.~word order). Taken together, our
results suggest a refinement of the Uniform Information Density
hypothesis: speakers may structure their utterances to optimize
information density, but they must do so under the predictable
constraints of their language.

\hypertarget{calculating-entropy-curves}{%
\subsection{Calculating entropy
curves}\label{calculating-entropy-curves}}

In all of our studies, we used an adaptation of the by-word entropy
method developed by Yu et al. (2016). Given a text or speech corpus
divided into individual sentences, we partition the corpus by sentence
length in number of words. For each word position \(X\) of sentences of
length \(k\), we define \(w\) as a unique word occurring in position
\(X\). We further define \(p(w)\) as the number of times word \(w\)
occurs in position \(X\), divided by the number of total words that
occur in position \(X\) i.e.~the number of sentences of length \(k\).
This creates a probability distribution over the words occurring in
position \(X\), and computing the Shannon (1948) entropy \(H(X)\) of
this probability distribution gives the positional entropy of position
\(X\) in sentences of length \(k\).

\[H(X) = \sum\limits_w p(w)\log\big(p(w)\big)\]

With this measure, we compute the unigram entropy at each word position
of sentences of each length within the corpus. The result of this method
can be plotted for each length as an \emph{entropy curve}, which can be
visually compared across sentence lengths to observe the how the unigram
entropy changes across absolute positions in each of the sentences.
Genzel \& Charniak (2002) similarly examine a unigram entropy measure on
sentences, and found that entropy at the sentence level increases
linearly with sentence index within a corpus. Uniform Information
Density applies this uniformity of entropy rate in sentences to all
levels of speech, and so our method obtained from Yu et al. (2016),
which examines text at the word level, should find a monotonically
increasing affine function at the word level.

The entropy curves capture individual variation across positions in
utterances of the same length. This allows us to directly observe and
judge the amount of variation in words that appear in an individual
position of a sentence. We can directly compare any two positions within
utterances to determine the amount of uncertainty, and therefore
information, on average contained by words within that position of
utterances. This method is thus identical in logic to Genzel \& Charniak
(2002), but within sentences instead of across sentences.

\hypertarget{study-1}{%
\section{Study 1}\label{study-1}}

We began by replicating Yu et al.'s (2016) analysis, computing entropy
curves on the British National Corpus--a collection of predominantly
written English (Clear, 1993). We then applied the same method to the
Switchboard corpus--a collection of spoken language (Godfrey, Holliman,
\& McDaniel, 1992). This allows us to ask whether the characteristic
function identified by Yu et al. (2016) is a general feature of English,
or instead a function of written language.

\hypertarget{data-and-analysis}{%
\section{Data and Analysis}\label{data-and-analysis}}

The British National Corpus consists of predominantly (90\%) written
language documents collected in the early 1980s and 1990s across a
variety of genres (scientific articles, newspapers, fiction, etc). It
also contains a small collection of corpora from spoken language. All
together, it contains \(\sim\) 100 million tokens. The Switchboard
corpus is a collection of \(\sim\) 2,400 telephone conversations between
unacquainted adults prompted with subjects of conversation. Switchboard
is the corpus used in Levy and Jaeger's (2007) original demonstration of
the Uniform Information Density Hypothesis.

For each corpus, we computed entropy curves using the method described
above for all sentences from length 4 to 30.

\hypertarget{results}{%
\section{Results}\label{results}}

The entropy curves computed in both the British National Corpus and
Switchboard were remarkably consistent both across corpora and across
the range of sentence lengths we analyzed (Figure
\ref{fig:read_and_plot_exp1}). Confirming Yu et al.'s (2016) findings,
we find that the positional entropy at the beginning of sentences is
low, then rises and plateaus for sentence-medial word positions before
dropping slightly in the second-to-last position and rising again.

\begin{CodeChunk}
\begin{figure*}[tb]

{\centering \includegraphics{figs/read_and_plot_exp1-1}

}

\caption[Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom)]{Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom). Points average entropies, error bars show 95\% confidence intervals computed by non-parametric bootstrap.}\label{fig:read_and_plot_exp1}
\end{figure*}
\end{CodeChunk}

The shape of positional entropy curves we find in these two corpora is
notable for two reasons: (1) it does not follow our predictions from
Uniform Information Density, and (2) the distribution is robust across
written and spoken Engilsh. This suggests that the entropy curve is
characteristic of the English language. We next asked whether this shape
is a feature even of children's speech, and also whether it varies
cross-linguistically.

\hypertarget{study-2}{%
\section{Study 2}\label{study-2}}

To understand how robust this entropy curve is, we turned to
conversational speech from parent-child interactions. If we find the
same shape even in children's productions, we have even stronger
evidence that the 3-step curve is a charcateristic feature of English.
We thus turned to the Child Language Data Exchange System (CHILDES), a
collection of transcripts of parent-child interactions (MacWhinney,
2014).

\hypertarget{data-and-analysis-1}{%
\subsection{Data and Analysis}\label{data-and-analysis-1}}

We analyzed three seperate corpora in CHILDES: The Providence Corpus
(Demuth, Culbertson, \& Alter, 2006), The Shiro corpus (Shiro, 2000),
and the Zhou Dinner Corpus (Li \& Zhou, 2015). The Providence corpus
consists of conversations between six 1--3-year-old American
English-speaking children and their parents recorded in their homes. The
Shiro Corpus consists of prompted Spanish-language narratives
individually collected from over a hundred Venezualan schoolchildren,
half from high SES backgrounds and half from low SES backgrounds. The
Zhou Dinner Corpus contains dinner conversations between 5--6-year-old
Mandarin-speaking children and their parents collected in Shanghai.
Spanish is an Indo-European language like English, possessing similar
grammar, word order and numerous cognate words. In contrast, Mandarin
Chinese is typologically unrelated to English.

We accessed the transcripts from each corpus using \texttt{childesr}, an
R-interface to a database formatted version of CHILDES (Sanchez et al.,
in press). We divided all utterances from each transcript into those
produced by target child, and those produced by all other speakers. We
then applied the same entropy curve method described above. For
Mandarin, we used pinyin transliterations of the utterances in the
corpus with demarcated word boundaries. The Chinese characters used for
writing Mandarin do not normally demarcate word boundaries by spacing
words apart, and for normal Chinese writing including spaces between
word boundaries is unnatural and can have a negative effect on reading
times (Bai, Yan, Liversedge, Zang, \& Rayner, 2008).

\hypertarget{results-1}{%
\subsection{Results}\label{results-1}}

\begin{CodeChunk}
\begin{figure*}[tb]

{\centering \includegraphics{figs/plot_childes-1}

}

\caption[Representative Entropy Curves for Three Childes Corpora in English, Mandarin, and Spanish]{Representative Entropy Curves for Three Childes Corpora in English, Mandarin, and Spanish. Points average entropies, error bars show 95\% confidence intervals computed by non-parametric bootstrap.}\label{fig:plot_childes}
\end{figure*}
\end{CodeChunk}

Across corpora, we found similar entropy curve shapes for adults and
children, but distinct shapes for each language. Figure
\ref{fig:plot_childes} shows representative curves across corpora, but
shapes were robust across the full range utterances we analyzed. We
found a distinct three-step distribution for English and Spanish CHILDES
corpora, with a slight dip in the penultimate position of each sentence.
The Mandarin corpus entropy curve, by comparison, has a noticeably lower
positional entropy values in utterance-final positions than in
utterance-penultimate positions.

These results present two important pieces of information. First, the
Providence corpus entropy curve broadly replicates the curves found in
both the British National Corpus and in Switchboard. Thus, the entropy
curve of English appears not just in adult-adult conversation, but even
in speech produced by parents to their children, and speech produced by
very young children to their parents. This suggests that it is a highly
robust feature of English language, as it structures the productions of
even pre-schooled aged children.

Second, the entropy curves of English, Spanish, and Mandarin were not
identical. No shape resembled the affine function predicted naively from
Uniform Information Density, but also Mandarin was quite different from
both English and Spanish. This suggests that the entropy curve can vary
from language to language. One possibility is that this variation arises
from typological features of languages, such as characteristic word
order. We noticed, for instance, a relatively high density of
determiners in the penultimate position of English and Spanish
utterances, which could account their characteristic penultimate dip. In
our final analysis, we explored this possibility directly, analyzing a
large set of Wikipedia corpora from diverse languages, and asking
whether variability in their entropy curves was related to variability
in their syntactic structure.

\hypertarget{study-3}{%
\section{Study 3}\label{study-3}}

In order to understand how entropy curves are related to linguistic
structure, we turned to Wikipedia for a larger and more diverse of set
of languages.

\hypertarget{method}{%
\subsection{Method}\label{method}}

\hypertarget{data-and-analysis-2}{%
\subsubsection{Data and Analysis}\label{data-and-analysis-2}}

Wikipedia has two primary advantages: (1) Individual language corpora
are generally large, and (2) hundreds of languages are represented. To
understand how variability in entropy curves is related to the structure
of language, we used features from the World Atlas of Language
Structures (WALS, Dryer \& Haspelmath, 2013). WALS is The World Atlas of
Language Structures is a curated collection of typological features
across the world's languages. However, as the collection includes
contributions from many authors studying a heterogeneous set of topic
and languages, most features are missing for most languages. For our
sample, we sought to balance diversity of languages with coverage in
WALS. We selected \(45\) languages for which we \(8\) WALS features were
coded (Figure \ref{fig:clust_tree}). Unfortunately, these features all
reflected the order of words in a sentence in each langauge
(e.g.~whether objects come before verbs, whether adjectives come before
nouns). A consistent set of results across these features would be
consistent with an effect of order on entropy curves, but unfortunately
no other feature types are available for comparison.

We computed entropy curves separately for each language as before. In
order to understand how these curves varied across languages, we
developed a method for compressing the curves into a few representative
values. We noticed that curves tended to vary in predictable utterance
positions: The majority were flat in the middle of utterances, but had
either rising or falling slopes at the beginnings and ends. For each
language, for each utterance length \(k\) we computed 5 slopes: The
change between the 1st and 2nd position, the change between the 2nd and
3rd position, the slope between the 3rd position and the 3rd-to-last
position (\(k-1\)), the slope between the 3rd-to-last and the
2nd-to-last position (\(k-1\)), and the slope between the 2nd-to-last
and final positions. We then averaged these slopes across utterance
lengths within a language to estimate a language-typical signature. We
performed an additional analysis dividing each utterance into 6ths
evenly and computing the slopes between these 6ths. Results from this
analysis were qualitatively similar.

\hypertarget{results-and-analysis}{%
\subsection{Results and Analysis}\label{results-and-analysis}}

We treated these 5 slopes as a vector, allowing us to compute the
pairwise similarity between languages with cosine--a standard measure of
vector similarity (Landauer \& Dumais, 1997). Intuitively, vectors with
with a smaller angle between them are more similar. Figure
\ref{fig:clust_tree} shows a hierarchical clustering of these 5
dimensional vectors, demonstrating that entropy curve similarity appears
to capture typological relatedly, e.g.~Slavic languages are similar to
each-other.

We then used asked whether these pairwise similarities are related to
overlap in linguistic features. For each of the 8 features we measured,
we fit a logistic regression predicting whether two languages would have
the same value for it from their pairwise entropy curve similarity.
Overlap on all 8 features was reliably predicted by entropy curve
similarity (largest \(p = .001\)), with the amount of variance in
feature overlap predicted varying across predictors
\ref{tab:tab_features}. Order of object and verb (\(83A\)) was the
strongest predictor, explaining 16.22\% of the variance in pairwise
entropy curve similarity.

\begin{CodeChunk}
\begin{figure}[tb]

{\centering \includegraphics{figs/clust_tree-1}

}

\caption[A dendrogram estimated from hierarchically clustering the slopes of our 45 languages]{A dendrogram estimated from hierarchically clustering the slopes of our 45 languages}\label{fig:clust_tree}
\end{figure}
\end{CodeChunk}

\begin{table}[tb]
\centering
\begin{tabular}{lrrr}
  \hline
\# & t-value & p-value & r2 \\ 
  \hline
83A & 178.22 & 0.00 & 0.16 \\ 
  95A & 170.98 & 0.00 & 0.15 \\ 
  81A & 153.70 & 0.00 & 0.13 \\ 
  97A & 136.81 & 0.00 & 0.10 \\ 
  144A & 74.43 & 0.00 & 0.03 \\ 
  87A & 39.91 & 0.00 & 0.01 \\ 
  143A & 39.74 & 0.00 & 0.01 \\ 
  82A & 3.23 & 0.00 & 0.00 \\ 
   \hline
\end{tabular}
\caption{Capptoin here} 
\label{tab:tab_features}
\end{table}

\hypertarget{general-discussion}{%
\section{General Discussion}\label{general-discussion}}

Taken together, the results of our studies yield 3 main findings. First,
languages have characteristic entropy curves that can be found in their
productions whether written, spoken, or even spoken by children. These
curves diverge in reliabile ways from the monotonically increasing
functions predicting by a naive Uniform Information Density hypothesis.
Second, these characteristic curves vary across languages--some
languages like English have increasing curves, but others like Mandarin
are characterized by decreasing entropy of the course of an utterance.
Finally, these characteristic curves are related to structural
properties of those languages, for instance their word orders. This work
presents an important rejoinder to models of linguistic productions as
optimizing communicative success. While speakers may do their best to
structure the information in their productions to minimize spikes in
entropy, they operate under coding models they inherit from their native
language, and these coding models constrain the choices they make.

\hypertarget{references}{%
\section{References}\label{references}}

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}

\noindent

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-anderson1991}{}%
Anderson, J. R. (1991). The adaptive nature of human categorization.
\emph{Psychological Review}, \emph{98}(3), 409.

\leavevmode\hypertarget{ref-bai2008}{}%
Bai, X., Yan, G., Liversedge, S. P., Zang, C., \& Rayner, K. (2008).
Reading spaced and unspaced chinese text: Evidence from eye movements.
\emph{Journal of Experimental Psychology: Human Perception and
Performance}, \emph{34}(5), 1277.

\leavevmode\hypertarget{ref-clear1993}{}%
Clear, J. H. (1993). The british national corpus. \emph{The Digital
World}, 163--187.

\leavevmode\hypertarget{ref-demuth2006}{}%
Demuth, K., Culbertson, J., \& Alter, J. (2006). Word-minimality,
epenthesis and coda licensing in the early acquisition of english.
\emph{Language and Speech}, \emph{49}(2), 137--173.

\leavevmode\hypertarget{ref-dryer2013}{}%
Dryer, M. S., \& Haspelmath, M. (2013). \emph{The world atlas of
language structures online}. Max Planck Institute for Evolutionary
Anthropology.

\leavevmode\hypertarget{ref-ferrer-i-cancho2013}{}%
Ferrer-i-Cancho, R., Debowski, L., \& Prado Martin, F. M. del. (2013).
Constant conditional entropy and related hypotheses. \emph{Journal of
Statistical Mechanics: Theory and Experiment}, \emph{2013}(07), L07001.

\leavevmode\hypertarget{ref-frank2008}{}%
Frank, A. F., \& Jaeger, T. F. (2008). Speaking rationally: Uniform
information density as an optimal strategy for language production. In
\emph{Proceedings of the annual meeting of the cognitive science
society} (Vol. 30).

\leavevmode\hypertarget{ref-genzel2002}{}%
Genzel, D., \& Charniak, E. (2002). Entropy rate constancy in text. In
\emph{Proceedings of the 40th annual meeting on association for
computational linguistics} (pp. 199--206). Association for Computational
Linguistics.

\leavevmode\hypertarget{ref-godfrey1992}{}%
Godfrey, J. J., Holliman, E. C., \& McDaniel, J. (1992). SWITCHBOARD:
Telephone speech corpus for research and development. In
\emph{Acoustics, speech, and signal processing, 1992. ICASSP-92., 1992
ieee international conference on} (Vol. 1, pp. 517--520). IEEE.

\leavevmode\hypertarget{ref-jaeger2013}{}%
Jaeger, T. F., \& Snider, N. E. (2013). Alignment as a consequence of
expectation adaptation: Syntactic priming is affected by the prime's
prediction error given both prior and recent experience.
\emph{Cognition}, \emph{127}(1), 57--83.

\leavevmode\hypertarget{ref-jain2018}{}%
Jain, A., Singh, V., Ranjan, S., Rajkumar, R., \& Agarwal, S. (2018).
Uniform information density effects on syntactic choice in hindi. In
\emph{Proceedings of the workshop on linguistic complexity and natural
language processing} (pp. 38--48).

\leavevmode\hypertarget{ref-jelinek1976}{}%
Jelinek, F. (1976). Continuous speech recognition by statistical
methods. \emph{Proceedings of the IEEE}, \emph{64}(4), 532--556.

\leavevmode\hypertarget{ref-landauer1997}{}%
Landauer, T. K., \& Dumais, S. T. (1997). A solution to plato's problem:
The latent semantic analysis theory of acquisition, induction, and
representation of knowledge. \emph{Psychological Review}, \emph{104}(2),
211.

\leavevmode\hypertarget{ref-levy2007}{}%
Levy, R. P., \& Jaeger, T. F. (2007). Speakers optimize information
density through syntactic reduction. In \emph{Advances in neural
information processing systems} (pp. 849--856).

\leavevmode\hypertarget{ref-li2015}{}%
Li, H., \& Zhou, J. (2015). \emph{Study on dinner table talk of
preschool children family in shanghai} (Master's thesis). East China
Normal University.

\leavevmode\hypertarget{ref-macwhinney2014}{}%
MacWhinney, B. (2014). \emph{The childes project: Tools for analyzing
talk, volume ii: The database}. Psychology Press.

\leavevmode\hypertarget{ref-mahowald2013}{}%
Mahowald, K., Fedorenko, E., Piantadosi, S. T., \& Gibson, E. (2013).
Info/information theory: Speakers choose shorter words in predictive
contexts. \emph{Cognition}, \emph{126}(2), 313--318.

\leavevmode\hypertarget{ref-piantadosi2011}{}%
Piantadosi, S. T., Tily, H., \& Gibson, E. (2011). Word lengths are
optimized for efficient communication. \emph{Proceedings of the National
Academy of Sciences}, \emph{108}(9), 3526--3529.

\leavevmode\hypertarget{ref-piantadosi2012}{}%
Piantadosi, S. T., Tily, H., \& Gibson, E. (2012). The communicative
function of ambiguity in language. \emph{Cognition}, \emph{122}(3),
280--291.

\leavevmode\hypertarget{ref-sanchez2019}{}%
Sanchez, A., Meylan, S. C., Braginsky, M., MacDonald, K. E., Yurovsky,
D., \& Frank, M. (in press). Childes-db: A flexible and reproducible
interface to the child language data exchange system. \emph{Behavior
Research Methods}.

\leavevmode\hypertarget{ref-shannon1948}{}%
Shannon, C. E. (1948). A mathematical theory of communication.
\emph{Bell System Technical Journal}, \emph{27}(3), 379--423.

\leavevmode\hypertarget{ref-shiro2000}{}%
Shiro, M. (2000). Diferencias sociales en la construcción del" yo" y
del" otro": Expresiones evaluativas en la narrativa de niños caraqueños
en edad escolar. In \emph{Lengua, discurso, texto: I simposio
internacional de análisis del discurso} (pp. 1303--1318). Visor.

\leavevmode\hypertarget{ref-simons2018}{}%
Simons, G. F., \& Fenning, C. D. (Eds.). (2018). \emph{Ethnologue:
Languages of the world, 21st edition}. Dallas, Texas: SIL International.

\leavevmode\hypertarget{ref-xu2018}{}%
Xu, Y., \& Reitter, D. (2018). Information density converges in
dialogue: Towards an information-theoretic model. \emph{Cognition},
\emph{170}, 147--163.

\leavevmode\hypertarget{ref-yu2016}{}%
Yu, S., Cong, J., Liang, J., \& Liu, H. (2016). The distribution of
information content in english sentences. \emph{arXiv Preprint
arXiv:1609.07681}.

\leavevmode\hypertarget{ref-zhan2018}{}%
Zhan, M., \& Levy, R. (2018). Comparing theories of speaker choice using
a model of classifier production in mandarin chinese. In
\emph{Proceedings of the 2018 conference of the north american chapter
of the association for computational linguistics: Human language
technologies, volume 1 (long papers)} (Vol. 1, pp. 1997--2005).

\bibliographystyle{apacite}

\end{document}