### 2019-02-11 ngrams/prelim_ngrams.Rmd
more childes bigrams and trigrams

---
title: "Eye-tracking"
author: "Joe Klafka"
date: "2/8/2019"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidytext)
library(tidyboot)
library(entropy)
library(magrittr)
library(feather)
library(here)

library(childesr)

knitr::opts_chunk$set(echo = TRUE)
```

```{r corpora}
prov_utterances <- get_utterances(corpus = "Providence") %>% 
  filter(speaker_role == "Target_Child") %>%
  mutate(words = stem)

okayama_utterances <- get_utterances(corpus = "Okayama")  %>% 
  filter(speaker_role == "Target_Child") %>%
  rename(words = gloss)

zhou_utterances <- get_utterances(corpus = "ZhouDinner") %>%
  filter(speaker_role == "Target_Child") %>%
  rename(words = stem)

switchboard <- read_feather(here("../switchboard/switchboard.feather")) %>%
  rename(words = value)
```

```{r ngram entropy}
# unnest_tokens(word, value, token = stringr::str_split, pattern = " +")

# takes in a tbl of utterances the column with the actual utterances is 
# named words, along with an integer target_length and a ngram length N.
ngram_entropy <- function(utterances, target_length, N) {

 tokens <- utterances %>%
  mutate(length = str_count(words, pattern = "[ +]+") + 1) %>%
  filter(length == target_length) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(ngrams, words, token = "ngrams", n = N) %>%
  group_by(utterance_id) %>%
  mutate(gram_order = 1:n()) %>%
  group_by(gram_order, ngrams) %>%
  filter(gram_order <= target_length - (N-1))

 tokens %>%
  summarise(n = n()) %>%
  tidyboot(summary_function = function(x) x %>% 
        summarise(entropy = entropy(n, unit = "log2")),
       statistics_functions = function(x) x %>%
       summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
  mutate(length = target_length)
}
```

```{r childes}
## Providence
# unigrams
prov_unigrams <- map(2:10, ~ngram_entropy(prov_utterances, .x, 1)) %>% 
  bind_rows()

prov_unigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

# bigrams
prov_bigrams <- map(3:10, ~ngram_entropy(prov_utterances, .x, 2)) %>% 
  bind_rows()

prov_bigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

# trigrams
prov_trigrams <- map(4:10, ~ngram_entropy(prov_utterances, .x, 3)) %>% 
  bind_rows()

prov_trigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

## Zhou
# unigrams
zhou_unigrams <- map(2:10, ~ngram_entropy(zhou_utterances, .x, 1)) %>% 
  bind_rows()

zhou_unigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

# bigrams
zhou_bigrams <- map(3:10, ~ngram_entropy(zhou_utterances, .x, 2)) %>% 
  bind_rows()

zhou_bigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

# trigrams
zhou_trigrams <- map(4:10, ~ngram_entropy(zhou_utterances, .x, 3)) %>% 
  bind_rows()

zhou_trigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

## Okayama
# bigrams 
okayama_bigrams <- map(3:10, ~ngram_entropy(okayama_utterances, .x, 2)) %>% 
  bind_rows()

okayama_bigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

# trigrams
okayama_trigrams <- map(4:10, ~ngram_entropy(okayama_utterances, .x, 3)) %>% 
  bind_rows()

okayama_trigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)
```

```{r zhou and okayama}
# bigrams
prov_bigrams <- map(3:10, ~ngram_entropy(prov_utterances, .x, 2)) %>% 
  bind_rows()

prov_bigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

# trigrams
prov_trigrams <- map(4:10, ~ngram_entropy(prov_utterances, .x, 3)) %>% 
  bind_rows()

prov_trigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

```

```{r switchboard}
# unigrams to check against previous results
switchboard_unigrams <- map(2:10, ~ngram_entropy(switchboard, .x, 1)) %>% 
  bind_rows()

switchboard_unigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

# bigrams
switchboard_bigrams <- map(3:10, ~ngram_entropy(switchboard, .x, 2)) %>% 
  bind_rows()

switchboard_bigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

#trigrams
switchboard_trigrams <- map(4:10, ~ngram_entropy(switchboard, .x, 3)) %>% 
  bind_rows()

switchboard_trigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)
```