### 2019-03-04 wiki_entropy/wiki_clustering.Rmd
clustering issues

---
title: Wikipedia Text Mining and Entropy Analysis
author: Josef Klafka and Dan Yurovsky
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
---

```{r setup, include=FALSE}
library(tidyverse)
library(googlesheets)
library(lingtypology)
library(knitr)
library(missMDA)
library(cluster)
library(widyr)
library(lsa)
library(magrittr)
library(broom)
library(here)
library(feather)
library(FactoMineR)
library(feather)
library(janitor)

knitr::opts_chunk$set(echo = TRUE)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 16))
```

```{r imputation and MCA}
wiki_gs <- gs_title("Wikipedia_absolute")
wkpd <- gs_read(wiki_gs)
LANGUAGES <- wkpd$Language

FEATURES <- c(1:144)
for (i in c(1:length(FEATURES))) {
  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
} #get all WALS features in correct syntax

wf <- wals.feature(FEATURES)

fair_wf <- wf[rowSums(!is.na(wf)) > 50,] %>%
  filter(language %in% LANGUAGES) %>%
  select(-wals.code, -latitude, -longitude, -glottocode) %>%
  as_tibble() %>%
  clean_names()

# Phonology `1A`:`19A`
# Morphology `20A`:`29A`
# Nominal Categories `30A`:`57A`
# Nominal Syntax `58A`:64A`
# Verbal syntax `65A`:`80A`
# Word order `81A`:`97A`, `143A`, `144A`
# Simple clauses `98A`:`120A`
# Complex sentences `121A`:`128A`
# Lexicon `129A`:`138A`
# Sign languages `139A`:`140A`

wa <- fair_wf %>% 
  filter(language %in% LANGUAGES) %>%
  select(x1a:x19a, language) %>%
  group_by(language) %>%
  slice(1)

wa_langs <- wa %>%
  gather(measure, value, -language) %>%
  group_by(language) %>%
  summarise(value = mean(!is.na(value))) %>%
  arrange(desc(value)) %>%
  filter(value > .6) %>% 
  pull(language)

wa_sub <- wa %>%
  filter(language %in% wa_langs) %>%
  column_to_rownames("language")
 # mutate_each(~as.numeric(as.factor(.x)))

wa_out <- wa_sub %>%
  select(-x14a:-x17a, -x9a,-x10a) %>%
  slice(1:5) %>%
  MIMCA(ncp = 1)

wa_imputed <- wa_out[["res.MI"]] %>%
  map(~mutate(.x, language = wa_langs)) %>%
  bind_rows(.id = "boot") %>%
  group_by(language, boot) %>%
  gather(measure, value, -language, -boot) %>%
  group_by(language, measure) %>%
  summarise(value = mean(value, na.rm = T)) %>%
  spread(measure, value)
```

```{r word order}
wiki_gs <- gs_title("Wikipedia_absolute")
wkpd <- gs_read(wiki_gs) 
LANGUAGES <- wkpd$Language

FEATURES <- c(82:97)
for (i in c(1:length(FEATURES))) {
  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
} #get all word order WALS features in correct syntax

wf <- wals.feature(FEATURES)

wo_wf <- wf %>% 
  filter(language %in% LANGUAGES) %>%
  select(language, `82A`:`97A`) %>%
  select(-`84A`)

wo_df <- wo_wf[rowSums(is.na(wo_wf)) < 6,] %>%
  as_data_frame() %>%
  group_by(language) %>%
  slice(1) %>%
  gather(feature, value, `82A`:`97A`) %>%
  group_by(feature) %>%
  mutate(value = factor(value))

imputed_wo <- wo_df %>%
  select(language, feature, value) %>%
  ungroup() %>% 
  spread(feature, value) %>% 
  column_to_rownames("language") %>%
  sapply(as.factor) %>%
  MIMCA(ncp = 1)

imputed_wo <- as.data.frame(imputed_wo[[1]][[99]]) 
imputed_wo$language <- unique(wo_df$language)
wo_sub <- imputed_wo %>% group_by(language) %>%
  slice(1) %>%
  gather(feature, value, `82A`:`97A`) %>%
  group_by(feature) %>%
  mutate(value = as.numeric(as.factor(value)))

wf_pairwise <- expand.grid(language1 = wo_sub$language,
                           language2 = wo_sub$language) %>%
  left_join(wo_sub, by = c("language1" = "language")) %>%
  rename(value1 = value) %>%
  left_join(wo_sub, by = c("language2" = "language", "feature")) %>%
  rename(value2 = value) %>%
  mutate(same = value1 == value2) %>%
  select(-value1, -value2) %>%
  left_join(cosines, by = c("language1", "language2"))

models <- wf_pairwise %>%
  group_by(feature) %>%
  nest() %>%
  mutate(model = map(data, ~glm(same ~ cosine, 
                                family = "binomial", data = .)))

models %>%
  mutate(coeffs = map(model, tidy)) %>%
  select(-data, -model) %>%
  unnest() %>%
  filter(term == "cosine") %>%
  arrange(desc(statistic)) %>%
  write_feather(here("word_order_glm_absolute.feather"))
```

```{r hierarchical clustering}
NUM_CLUSTERS <- 10

##read in googlesheet
wiki_gs <- gs_title("Wikipedia_absolute")
wkpd <- gs_read(wiki_gs) 
LANGUAGES <- wkpd$Language

wkpd <- wkpd %>% 
  as.data.frame() %>%
  column_to_rownames(var = "Language")

##do clustering
wk_clust_tree <- wkpd %>% 
              dist() %>%
              hclust("ward.D2")

wk_clust_tree %>% plot(hang = -1)
wk_clusters <- wk_clust_tree %>%
              cutree(k = NUM_CLUSTERS)

wk_clusters %>% 
  t() %>%
  as_data_frame() %>% 
  t() %>%
  as_data_frame(rownames = "language") %>%
  rename(cluster = V1) %>%
  arrange(cluster) %>% 
  View()
```

```{r}
wiki_gs <- gs_title("Wikipedia_relative5")
wkpd <- gs_read(wiki_gs) 
LANGUAGES <- wkpd$Language

FEATURES <- c(1:144)
for (i in c(1:length(FEATURES))) {
  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
} #get all WALS features in correct syntax

wf <- wals.feature(FEATURES)

wf_df <- wf %>%
  as_data_frame() %>%
  filter(language %in% LANGUAGES) %>%
  group_by(language) %>%
  slice(1) %>%
  gather(feature, value, `1A`:`144A`) %>%
  group_by(feature) %>%
  mutate(value = factor(value))

cosines <- wkpd %>%
  column_to_rownames("Language") %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, cosine, -language1)

```

```{r cogsci glm subset}

wf_many_langs <- wf_df %>%
  group_by(feature) %>%
  summarise(n = sum(!is.na(value))) %>%
  filter(n > 60)

wf_sub_langs <- wf_df %>%
  filter(feature %in% wf_many_langs$feature) %>%
  group_by(language) %>%
  summarise(keep = sum(is.na(value)) == 0) %>%
  filter(keep)

wf_sub <- wf_df %>%
  filter(feature %in% wf_many_langs$feature,
         language %in% wf_sub_langs$language) %>%
  select(language, feature, value) %>%
  mutate(value = as.numeric(as.factor(value)))

wf_pairwise <- expand.grid(language1 = wf_sub$language,
                           language2 = wf_sub$language) %>%
  left_join(wf_sub, by = c("language1" = "language")) %>%
  rename(value1 = value) %>%
  left_join(wf_sub, by = c("language2" = "language", "feature")) %>%
  rename(value2 = value) %>%
  mutate(same = value1 == value2) %>%
  select(-value1, -value2) %>%
  left_join(cosines, by = c("language1", "language2"))

models <- wf_pairwise %>%
  group_by(feature) %>%
  nest() %>%
  mutate(model = map(data, ~glm(same ~ cosine, 
                                family = "binomial", data = .)))

models %>%
  mutate(coeffs = map(model, tidy),
         rs = map(model, )) %>%
  select(-data, -model) %>%
  unnest() %>%
  filter(term == "cosine") %>%
  arrange(desc(statistic))
```