### 2019-03-21 ngrams/wiki_ngrams.Rmd
starting wikipedia ngrams process

---
title: "Wikipedia ngrams Entropy Analysis"
author: "Josef Klafka and Dan Yurovsky"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidytext)
library(googlesheets)
library(here)
library(feather)

knitr::opts_chunk$set(echo = TRUE)
```

```{r getting gsheets}
wiki_gs <- gs_title("Wikipedia_relative5")
w1 <- gs_read(wiki_gs) %>% arrange(Language)
wiki_gs <- gs_title("Wikipedia_relative_2grams")
w2 <- gs_read(wiki_gs) %>% arrange(Language)

w1 %>% 
  filter(Language %in% w2$Language)

```