### 2019-04-04 ngrams/wiki_ngrams.Rmd
initial 1-4grams for wikipedia relative slopes

---
title: "Wikipedia ngrams Entropy Analysis"
author: "Josef Klafka and Dan Yurovsky"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidytext)
library(googlesheets)
library(here)
library(feather)
library(magrittr)
library(reshape2)

knitr::opts_chunk$set(echo = TRUE)
```

```{r getting gsheets}
# wiki_gs <- gs_title("Wikipedia_relative5")
# w1 <- gs_read(wiki_gs) %>% arrange(Language)
# wiki_gs <- gs_title("Wikipedia_relative_2grams")
# w2 <- gs_read(wiki_gs) %>% arrange(Language)
# wiki_gs <- gs_title("Wikipedia_relative_3grams")
# w3 <- gs_read(wiki_gs) %>% arrange(Language)
# wiki_gs <- gs_title("Wikipedia_relative_4grams")
# w4 <- gs_read(wiki_gs) %>% arrange(Language)

w1 <- read_feather(here("relative_unigrams.feather"))
w2 <- read_feather(here("relative_bigrams.feather"))
w3 <- read_feather(here("relative_trigrams.feather"))
w4 <- read_feather(here("relative_quadgrams.feather"))

get_agg <- function(ws) {
 ws %>%
  mutate(agg = abs(Slope1) + abs(Slope2) + abs(Slope3) + abs(Slope4) + abs(Slope5)) %>%
  select(Language, agg)
}

w1 %<>% 
  get_agg() %>%
  rename(unigram = agg)
w2 %<>%
  get_agg() %>%
  rename(bigram = agg)
w3 %<>%
  get_agg() %>%
  rename(trigram = agg)
w4 %<>%
  get_agg() %>%
  rename(quadgram = agg)

w1 %>% 
  left_join(w2) %>%
  left_join(w3) %>%
  left_join(w4) %>%
  remove_missing() %>%
  slice(1:5) %>%
  melt(id.vars = "Language") %>%
  ggplot(aes(variable, value, color=Language)) + 
  geom_point() +
  geom_line()

```