### 2019-04-18 ngrams/wiki_ngrams.Rmd
reorganized wikipedia ngrams data

---
title: "Wikipedia ngrams Entropy Analysis"
author: "Josef Klafka and Dan Yurovsky"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidytext)
library(here)
library(feather)
library(magrittr)
library(reshape2)
library(lme4)

knitr::opts_chunk$set(echo = TRUE)
```

```{r set up data}
w1 %>% 
  left_join(w2) %>%
  left_join(w3) %>%
  left_join(w4) %>%
  remove_missing() %>%
  slice(1:5) %>%
  melt(id.vars = "Language") %>%
  ggplot(aes(variable, value, color=Language)) + 
  geom_point() +
  geom_line()
```

```{r Joe's regression}
# log(sum(abs(s_i))) = language * type * gram + (source | type)

# let's just try log(sum(abs(s_i))) = language + gram since type/source are all the same atm
## why would there be an interaction between language and gram?
wn <- read_feather(here("Data/wiki_ngrams.feather"))

simple_glm <- glm(data = wn, formula = s_i ~ gram + Language)

```

```{r get data}
w1 <- read_feather(here("Data/relative_unigrams.feather"))
w2 <- read_feather(here("Data/relative_bigrams.feather"))
w3 <- read_feather(here("Data/relative_trigrams.feather"))
w4 <- read_feather(here("Data/relative_quadgrams.feather"))

get_agg <- function(ws) {
 ws %>%
  mutate(agg = abs(Slope1) + abs(Slope2) + abs(Slope3) + abs(Slope4) + abs(Slope5)) %>%
  select(Language, agg)
}

w1 %<>% 
  get_agg() %>%
  rename(unigram = agg)
w2 %<>%
  get_agg() %>%
  rename(bigram = agg)
w3 %<>%
  get_agg() %>%
  rename(trigram = agg)
w4 %<>%
  get_agg() %>%
  rename(quadgram = agg)

wn <- w1 %>% 
  left_join(w2) %>%
  left_join(w3) %>%
  left_join(w4) %>%
  remove_missing() %>%
  melt(id.vars = "Language") %>% 
  rename(gram = variable, s_i = value)

wn %>% write_feather(here("Data/wiki_ngrams.feather"))
```