### 2019-05-14 surprisal/childes_surprisal.Rmd
more changes

---
title: Positional surprisal (1-3grams) in CHILDES corpora
author: Josef Klafka and Dan Yurovsky 
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
# load packages
library(knitr)
library(tidyverse)
library(directlabels)
library(childesr) #data
library(tidytext)
library(entropy)
library(tidyboot)
library(dplyr)
library(tokenizers)
library(gtools)
library(here)
library(ggthemes)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 16))
```

Get utterances
```{r get_utterances}
prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
zhou_utterances <- get_utterances(corpus = "ZhouDinner") #stem for pinyin
shiro_utterances <- get_utterances(corpus = "Shiro")
```

```{r old unigrams}
adult_utterances <- shiro_utterances %>%
  filter(speaker_role == "Target_Child") %>% 
  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n())

adult_unigrams <- adult_utterances %>%
  group_by(word) %>%
  count() %>%
  ungroup() %>%
  mutate(p = n / sum(n))

adult_surprisals <- adult_utterances %>%
  left_join(adult_unigrams) %>%
  mutate(s = -log(p)) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s))

adult_surprisals %>%
  filter(length %in% c(5, 7, 9)) %>%
  ggplot(aes(x = word_order, y = s)) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_smooth(se = F)
```

```{r unigrams}
get_unigrams <- function(utterances, speaker_role, stem_or_gloss) {

  if (tolower(speaker_role) == "child") {
    utts <- utterances %>%
      filter(speaker_role == "Target_Child") 
  } else {
    utts <- utterances %>%
      filter(speaker_role != "Target_Child")   
  }

  if (tolower(stem_or_gloss) == "stem") {
    utts <- utts %>%
      mutate(length = str_count(stem, pattern = "[ +]+") + 1) %>%
      mutate(utterance_id = 1:n()) %>%
      unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+") %>%
      group_by(utterance_id) %>%
      mutate(word_order = 1:n())

  } else {
    utts <- utts %>%
      mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
      mutate(utterance_id = 1:n()) %>%
      unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
      group_by(utterance_id) %>%
      mutate(word_order = 1:n())
  }

  unigrams <- utts %>%
    group_by(word) %>%
    count() %>%
    ungroup() %>%
    mutate(p = n / sum(n)) %>%
    mutate(s = -log(p)) %>%
    select(word, s)

  surprisals <- utts %>%
    left_join(unigrams) %>%
    group_by(length, word_order) %>%
    summarise(s = mean(s)) %>%
    ungroup()
}

prov_sups <- get_unigrams(prov_utterances, "not child", "stem") %>% 
  mutate(speaker_role = "adult") %>%
  bind_rows(mutate(get_unigrams(prov_utterances, "child", "stem"), speaker_role = "child")) %>%
  mutate(language = "English")

shiro_sups <- get_unigrams(shiro_utterances, "not child", "gloss") %>% 
  mutate(speaker_role = "adult") %>%
  bind_rows(mutate(get_unigrams(shiro_utterances, "child", "gloss"), speaker_role = "child")) %>%
  mutate(language = "Spanish")

zhou_sups <- get_unigrams(zhou_utterances, "not child", "stem") %>% 
  mutate(speaker_role = "adult") %>%
  bind_rows(mutate(get_unigrams(zhou_utterances, "child", "stem"), speaker_role = "child")) %>%
  mutate(language = "Mandarin")

all_sups <- prov_sups %>%
  bind_rows(shiro_sups, zhou_sups)

label_data <- data_frame(length = 9, Language = "Spanish", word_order = 3,
                         speaker_role = c("Adult", "Child"), s = c(1.25,3.5))

all_trigrams %>% 
  filter(length %in% c(5, 7, 9)) %>%
  ggplot(aes(x = word_order, y = s,
                        color = speaker_role,
                        label = speaker_role)) +
    facet_grid(Language ~ length, scales = "free") + 
    xlab("Word position") +
    ylab("Mean surprisal") +
    geom_line(size = 1) +
    geom_point() + 
    theme_classic(base_size = 10) + 
    scale_color_ptol() +
    theme(legend.position = "nones") + 
    geom_text(data = label_data)

```

```{r bigrams}
adult_bigrams <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  group_by(lag_word, word) %>%
  count() %>%
  filter(!is.na(lag_word)) %>%
  ungroup() %>%
  mutate(joint_p = n / sum(n)) %>%
  select(-n) %>%
  left_join(adult_unigrams, by = c("lag_word" = "word")) %>%
  mutate(cond_p = joint_p / p) %>%
  select(-n, -p)

bigram_surprisals_prep <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  left_join(adult_bigrams) %>%
  left_join(adult_unigrams)

bigram_surprisals_nonnull <- bigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)

all_bigrams <- all_bigrams %>% 
  bind_rows(mutate(bigram_surprisals_nonnull, speaker_role = "Child", Language = "Spanish"))

bigram_surprisals_nonnull %>%
  filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%
  ggplot(aes(x = word_order, y = s)) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_smooth(se = F)
```

```{r trigrams}
adult_trigrams <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  group_by(lag_word2, lag_word1, word) %>%
  count() %>%
  filter(!is.na(lag_word1), !is.na(lag_word2)) %>%
  ungroup() %>%
  mutate(tri_joint_p = n / sum(n)) %>%
  select(-n) %>%
  left_join(adult_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
  mutate(tri_cond_p = tri_joint_p / joint_p) %>%
  select(-joint_p, -tri_joint_p, -cond_p) # %>%

trigram_surprisals_prep <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  left_join(adult_trigrams) %>% 
  left_join(select(adult_bigrams, -joint_p), by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
  left_join(adult_unigrams)

trigram_surprisals_nonnull <- trigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
                    ifelse(is.na(lag_word1), # check if not the second word
                           -log(p), # trigram if third or beyond
                           -log(cond_p)), # bigram if second word
                    -log(tri_cond_p))) %>% # unigram if first word
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)

all_trigrams <- all_trigrams %>% 
  bind_rows(mutate(trigram_surprisals_nonnull, speaker_role = "Child", Language = "Spanish"))

trigram_surprisals_nonnull %>%
  filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%
  ggplot(aes(x = word_order, y = s)) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_smooth(se = F)

```