### 2019-05-28 surprisal/MacWhinney poster/MacWhinney poster.Rmd
posterdown not working yet

---
# PLEASE SEE THE README for in depth description github.com/brentthorne/posterdown
poster_height: "32in"
poster_width: "46in"
font_family: 'Rasa'
#ESSENTIALS
title: A Better Reproducible Poster Title
author:
  - name: W. Brent Thorne
    affil: 1,*
    main: true
    orcid: '0000-0002-1099-3857'
    twitter: brentthorne18 
    email: bthorne2@brocku.ca
  - name: Another G. Contributor
    affil: 2
    email: ag.con@posterdown.net
  - name: Person Three
    affil: 3
affiliation:
  - num: 1
    address: Department of Earth Science, Brock University
  - num: 2
    address: Department of Graphics and Layouts, University of Posters; Canada
  - num: 3
    address: Another Institute of a place where work gets done, Earth, Milky Way
#STYLE & FORMATTING
title_textsize: "125pt"
author_textsize: "1.17em"
authorextra_textsize: "35px"
affiliation_textsize: "25px"
affiliation_textcol: '#00000060'
caption_fontsize: "20pt"
#Middle of the poster
middle_fontfamily: "Special Elite"
middle_textcol: "#FFFFFF90"
middle_fontsize: "170px"
main_findings:
  - "Make **better posters** with <br> RMarkdown + **posterdown**."
  - "Transition from **poster** to **manuscript** with ease!"
logoleft_name: "Figures/qr-code-black.png"
logoright_name: "Figures/posterdownlogo.png"
#---POSTER BODY OPTIONS---#
primary_colour: '#0b4545'
secondary_colour: '#008080' 
accent_colour: "#cc0000"
body_bgcol: "#FFFFFF"
body_textsize: "45px"
body_textcol: "#000000"
#--Standard Options--#
output: 
  posterdown::posterdown_betterland:
    self_contained: false
    pandoc_args: --mathjax
    highlight: espresso
    number_sections: false
link-citations: true
bibliography: MyBib.bib
---

```{r, include=FALSE}
knitr::opts_chunk$set(results = 'asis', echo = FALSE, comment = FALSE,  warning = FALSE,tidy = FALSE, message = FALSE, fig.align = 'center')
options(knitr.table.format = "html") 
```

# Introduction

This is the `posterdown_betterland` template for the {posterdown} package! I was inspired by the twitter thread of [Mike Morrison](https://mobile.twitter.com/mikemorrison/status/1110191245035479041) and wanted to apply the `#betterposter` concept to the reproducible (yet simple to use) functionality of the {posterdown} package [@thorneposterdown2019]. If you're not an R user don't sweat as you do **NOT** need to use it at all! Feel free to use only the Markdown functionality of this package :)

# Methods

1. Install R and `posterdown`, see [github](github.com/brentthorne/posterdown)
2. Open the `posterdown_betterland` template
3. Make all your poster dreams come true!

# Results

```{r, out.width='90%', fig.height=5.5, fig.cap='Using {ggplot} and {patchwork} to generate a layout of multiple plots in one figure [@patchwork2017].'}

library(tidyverse)
library(patchwork)
library(ggthemes)

theme_set(theme_tufte() + theme(legend.position = "none"))

base <- ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, colour = Species))

p_point <- base + geom_point()
p_line  <- base + geom_line() 
p_area  <- base + geom_area()
p_box1  <- ggplot(iris) + geom_boxplot(aes(x = Species, y = Sepal.Length, fill = Species))
p_box2  <- ggplot(iris) + geom_boxplot(aes(x = Species, y = Sepal.Width, fill = Species))

p_line +
  {p_point + p_box1} +
  plot_layout(ncol = 1) + plot_annotation(tag_levels = "a", tag_prefix = "(",tag_suffix = ")")

```

## More Figures and Tables

```{r, mytable, fig.align='center', out.width='90%'}

library(knitr)
library(kableExtra)

dt <- iris[1:5, 1:5]

kable(dt, caption = 'Here is a caption for the table made with the {kableExtra} package [@kableExtra2019].',align = 'c', col.names = c("Sepal W","Sepal L","Pedal W","Pedal L","Species")) %>% 
  kable_styling(full_width = T) %>% 
  row_spec(0, color = "#fff", background = "#0b454580")
```

```{r, echo=TRUE}
library(ggplot2)
library(ggspatial)

load_longlake_data()

map <- ggplot() + 
  theme_pander() +
  layer_spatial(longlake_depthdf,
                aes(colour = DEPTH)) +
  theme(legend.position = "bottom")
```

```{r, mymapfig, out.width='70%', cache=FALSE, fig.cap='Map of Long Lake eample from the [ggspatial](https://github.com/paleolimbot/ggspatial) package [@ggspatial].',strip.white= TRUE, comment=FALSE}

ggsave(map,filename = "Figures/map.svg", width = 3.5, height = 3.5, device = 'svg', bg = 'transparent')

knitr::include_graphics("Figures/map.svg")
```

A **BIG** thank you to [Romain Leasur](https://twitter.com/RLesur) and [Yihui Xie](https://twitter.com/xieyihui) for their wonderful work on [{pagedown}](https://github.com/rstudio/pagedown) which had made this poster possible [@pagedown]!

# References