### 2019-05-28 surprisal/MacWhinney poster/MacWhinney_Poster.Rmd

---
# PLEASE SEE THE README for in depth description github.com/brentthorne/posterdown
#
#---POSTER SIZE & DEFAULT FONT---#
poster_height: "32in" # height in inches of poster
poster_width: "40in" # width in inches of poster
font_family: "palatino" # choose from typical latex fonts (example: "palatino")
font_size: "30pt" #please see github.com/brentthorne/posterdown for compatible options.

#---TITLE BOX OPTIONS---#
#ESSENTIALS
title: '\fontfamily{phv}\selectfont Parents and children in diverse languages structure their utterances to support efficient communication'
author: "Josef Klafka Daniel Yurovsky"
affiliation: "^1^Department of Psychology, University of Chicago"
#STYLE & FORMATTING
titlebox_bgcol: "008080"  #Colour of the Title Box background
titlebox_bordercol: "0b4545" #Colour of the title Box border.
titlebox_shape: "all"
titlebox_borderwidth: "1cm"
title_textcol: "ffffff" #colour of title text
author_textcol: "0b4545" # Colour of author text
affiliation_textcol: "FFFFFF" # Colour of affiliation text
title_textsize: "Huge"         # Poster title fontsize
author_textsize: "Large"       # Author list font size
affiliation_textsize: "large"  # Affiliation font size
#ADDING LOGOS
logoleft_name: 'Figures/posterdownlogo'
logoleft_width: '3in'
logoleft_xshift: '1in'
logoleft_yshift: '1in'
logoright_name: 'Figures/posterdownlogo'
logoright_width: '3in'
logoright_xshift: '-1in'
logoright_yshift: '1in'

#---POSTER BODY OPTIONS---#
body_bgcol: "ffffff" #colour of the poster main background
body_textsize: "normalsize"    # Size of the main poster body text
body_textcol: "000000" # Colour of main text in the body of poster
column_numbers: 2 # Number of columns that the poster has
column_margins: "0.5in" # Margin spacing for columns
columnline_col: "008080" #colour 
columnline_width: "0pt" #width of line between each column
#SECTION TITLE STYLING
sectitle_textcol: "ffffff" # Colour of the poster section titles
sectitle_bgcol: "0b4545" # Colour of the section title box
sectitle_bordercol: "0b4545" # Colour of the border around the section title box.
sectitle_borderwidth: "2mm" # Thicknes of the section title box border
sectitle_boxshape: "uphill" # Changes the shape of the section title box.

#---BIBLIOGRAPHY OPTIONS---#
bibliography: MyLibrary # name of the .bib file used for referencing
bibliography_spacing: 0.8 # sets the multiplier for line spacing of bibliography spacing (between 0 and 1)
bibliography_textsize: "small"  # size of the bibliography text size (handy for one too many references!)

#---OTHER---#
cite_col: "CC0000" #colour of ciation elements
url_col: "008080" #colour of url links
link_col: "008080" #colour of other links within the poster
footnote_textcol: "ffffff" # Colour of footnote text if used
output: posterdown::posterdown_latex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  results = 'asis',
  echo = FALSE,
  comment = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = 'center'
)
```

# Introduction

Welcome to `posterdown` ! This is my attempt to provide a semi-smooth workflow for those who wish to take their R Markdown skills to the conference world. Most features from R Markdown are available in this package such as Markdown section notation, figure captioning, and even citations like this one [@rmarkdown]. The rest of this example poster will show how you can insert typical conference poster features into your own document.

## Objectives

1. Easy to use reproducible poster design. 
2. Integration with R Markdown.
3. Easy transition from `posterdown` to `thesisdown` or `rticles` [@rticles; @thesisdown].

# Methods

This package uses the same workflow approach as the R Markdown you know and love. Basically it goes from RMarkdown > Knitr > Markdown > Pandoc > HTML/CSS > PDF. You can even use the bibliography the same way [@turnerControlsWaterBalance2014].

## Random text

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam placerat augue at velit tincidunt semper. Donec elementum porta posuere. Nullam interdum, odio at tincidunt feugiat, turpis nisi blandit eros, eu posuere risus felis non quam. Nam eget lorem odio. Duis et aliquet orci. Phasellus nec viverra est. Praesent odio orci, mattis vel mauris nec, consectetur fermentum mauris. Etiam eu hendrerit tortor. Donec mi tellus, efficitur et porttitor eu, auctor eu tellus. Quisque faucibus vestibulum sapien vel lacinia. Ut auctor lorem non interdum blandit.

# Results

Usually you want to have a nice table displaying some important results that you have calculated. In `posterdown` this is as easy as using the `kable` table formatting you are probably use to as per typical R Markdown formatting. I suggesting checking out the `kableExtra` package and its in depth documentation on customizing these tables found [here](https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf) [@kableExtra2019]. You can reference tables like so: Table \@ref(tab:mytable).

```{r mytable, out.width='80%'}
knitr::kable(iris[1:5, 1:4], caption = 'Table caption.',align = 'c',"html")
```

Or with figures: Figure \@ref(fig:standard-plot).

```{r standard-plot, out.width='80%', fig.align='center', echo=TRUE, fig.cap='Great figure!'}
plot(mtcars[1:2])
```

# Next Steps: More random text

Aliquam sed faucibus risus, quis efficitur erat. Vestibulum semper mauris quis tempus eleifend. Aliquam sagittis dictum ipsum, quis viverra ligula eleifend ut. Curabitur sagittis vitae arcu eget faucibus. In non elementum felis. Duis et aliquam nunc. Nunc pulvinar sapien nunc, vel pretium nisi efficitur in. Fusce fringilla maximus leo et maximus. Fusce at ligula laoreet, iaculis mi at, auctor odio. Praesent sed elementum justo. Aenean consectetur risus rhoncus tincidunt efficitur. Praesent dictum mauris at diam maximus maximus [@thorneposterdown2019].

# Conclusion

Try `posterdown` out! Hopefully you like it!

# References

\printbibliography