### 2019-07-01 surprisal/phoneme_writeup.Rmd
tidying phoneme writeup

---
title: "Phoneme surprisal"
author: Josef Klafka and Dan Yurovsky 
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: hide 
editor_options: 
  chunk_output_type: console
---

```{r load-libraries, library, message=F, results='hide', warning=FALSE}
library(tidytext)
library(tidyverse)
library(childesr)
library(glue)
library(feather)
library(here)
library(entropy)
library(tidyboot)
library(janitor)
library(arrangements)
library(stringr)

knitr::opts_chunk$set(echo = TRUE, cache = T, warning = F, 
                      fig.width = 8, fig.height = 4)

theme_set(theme_classic(base_size = 16))
```

Where is the information in English words? A salient feature of African American Enlgish Vernacular is the reduction of final consonants and consonant clusters. Why must this happen? One possibility is that the information important for identifying a word is not uniformly distributed across it's characters or sounds but instead concentrated at the beginnings or middles of words. In these cases, reducing final sounds might reduce the articulatory complexity of words without much cost for the listener in processing these words and inferring their intended meaning.

In a series of initial analyses we tackle three ways:
1. Computing the surprisal of characters in English words either from baserates of character production (unigram) or conditioning on the prior charaters in each word (bigram and trigram). This analyses roughly captures the difficulty that a listener might have in predicting each character.
2. Repeating this analysis at the level of phonemes rather than characters, arguably a better proxy for the speaker's and listener's representations
3. Computing hold-out entropy for individual phonemes in English words. This is roughly equivalent to asking "if you knew every other sound in a word, how much trouble would you have predicting what sounds appears in position X"
4. Comparing English to French and German which have different expected information profiles (more below).

SOME KIND OF PREVIEW OF RESULTS HERE?

```{r read-data}
prov_letters <- read_feather(here("Data/providence_letters.feather")) %>%
  mutate(rep = "letter", language = "English") %>%
  rename(order = letter_order)

prov_phones <- read_feather(here("Data/providence_phonemes.feather")) %>%
  mutate(rep = "phone", language = "English") %>%
  rename(order = phone_order)

adult_phonemes <- read_feather(here("Data/adult_phonemes.feather"))

french_letters <- read_feather(here("Data/palasis_letters.feather")) %>%
  mutate(rep = "letter", language = "French") %>%
  rename(order = letter_order)

german_letters <- read_feather(here("Data/wagner_letters.feather")) %>%
  mutate(rep = "letter", language = "German") %>%
  rename(order = letter_order)

all_data <- bind_rows(prov_letters, prov_phones, french_letters, german_letters)
```

# English Letters

We use for our English corpus the Providence corpus from CHILDES. In the plots below, each plot represents the information in each letter in words of different lengths by number of letters--we show the length above each plot.

We see a distinctive shape for information at the level of individual letters in English. The first letter has relatively high information, the second letter has less information and the third letter has high infomration, and the final letter (if the word has more than 3 letters) has less information.

```{r plot-helper}
plot_curves <- function(select_language = "English", select_rep = "letter", 
                        select_gram = "Unigram") {
  all_data %>%
    filter(language == select_language, rep == select_rep,
           gram == select_gram) %>%
    filter(word_length %in% 2:6) %>%
    ggplot(aes(x = order, y = mean_s)) +
    facet_grid(. ~ word_length, 
               labeller = as_labeller(function(x) paste0("Length ", x))) +
    geom_point() +
    geom_line() +
    xlab(paste0(str_to_sentence(select_rep), " Position")) +
    ylab("Surprisal") + 
    ggtitle(paste0("Information in ", select_language, " ", 
                   select_rep, "s by ",select_rep, " position - ", select_gram))
}
```

```{r prov-letter-unigrams}
plot_curves()
```

Now we ask: if you know the previous letter, how much information is in the next letter in English? There is relatively less information in each letter knowing the previous letter, but we see the same general information trajectory across letters.

```{r prov-letter-bigrams}
plot_curves(select_gram = "Bigram")
```

What about if you know the last two letters and are trying to predict the next letter? We see decreasing information within words as you move left-to-right.

```{r prov-letter-trigrams}
plot_curves(select_gram = "Trigram")
```

# English Phonemes

## Surprisal

Briefly, what we're doing is splitting each word in the Providence corpus from CHILDES up into its constituent phones, i.e. the sounds that make up the word. We then calculate how common each phone is across the entire corpus. From this frequency measure, we compute the surprisal of each phone, which is inversely proportional to the frequency of the phone. The less common a phone is, the more difficult it is to process the phone.

We split up the words in the corpus by their length in number of phones--for example, the word "food" has three phones: one for the "f", one for the long "oo", and one for the "d". We would take all words like this that have three phones. For each phone position  within the words of that length, we calculate the average surprisal/information in that position. For example, looking at all of the words with three phones, we take all of the first phones and calculate how much information is in that position on average. In other words, how difficult is it to process the first sound in a word, compared to the other sounds in a word?

Unlike with letters, there isn't as much of a consistent shape to the information distribution for phones considered without context. For the most part, the first phone has relatively little information while the second phone has a lot of information.

```{r prov-phone-unigrams}
plot_curves(select_rep = "phone")
```

With bigrams and trigrams for English phones, the next phone is very predictable if you know the phone or two that came before it. Especially if you know the preceding two phones, then each phone past the first two is predictable and carries relatively little information in English.

```{r prov-phone-bigrams}
plot_curves(select_rep = "phone", select_gram = "Bigram")
```

```{r prov-phone-trigrams}
plot_curves(select_rep = "phone", select_gram = "Trigram")
```

## Hold-out entropy

For this next analysis, we took all words with 2, 3, 4, 5 or 6 phones. For each word length, we held out the phones in one phone position. For example, in all words with 4 phones, we held out the first phone. We then computed the entropy, or amount of variation, in the remaining words. The higher the entropy when holding out phones in a given position, the more information is contained in that phone on average.

```{r leave-out-entropy, include = F}
hold_out_entropy <- function(df) {
  spread_df <- df %>%
    spread(phone_order, phone) %>%
    clean_names()

  phone_length <- max(df$phone_order)

  lens <- paste0("x", 1:phone_length)

  boot_entropy <- function(group) {
    spread_df %>%
      group_by_at(group) %>%
      summarise(n = n()) %>%
      summarise(entropy = entropy(n)) %>%
      ungroup() %>%
      tidyboot_mean(entropy)
  }

  groupings <- combinations(lens, phone_length -1, 
                            layout = "list")

  groupings <- map(seq(length(lens),1,-1),
                   function(x) c(groupings[[x]], lens[x]))

  map_df(groupings, boot_entropy) %>%
    mutate(phone_order = seq(phone_length, 1, -1))

}
```

```{r show-leave-out-entropy}
split_lengths <- adult_phonemes %>%
  group_by(word_length) %>%
  nest() %>%
  filter(word_length > 1) %>%
  filter(word_length %in% 1:6 ) %>%
  mutate(entropy = map(data, hold_out_entropy))

split_lengths %>%
  select(-data) %>%
  unnest() %>%
  ggplot(aes(x = phone_order, y = empirical_stat)) + 
  facet_wrap(~ word_length, 
               labeller = as_labeller(function(x) paste0("Length ", x)),
               scales = "free") +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) + 
  theme_classic() + 
  geom_line() + 
  xlab("Phone position") + 
  ylab("Entropy")
```

We see that word-initial phones tend to have more information in them than word-medial or word-final phones, so less is lost by dropping the final sounds in a word than in dropping the initial sounds.

We are interested in French because when many French words are pronounced, the final letters are not pronounced in any way (e.g. "ils sauraient"="they would know", the final "ent" in "sauraient" is not pronounced at all). We are interested in German because German has a much richer case system than English or French, and case markings in German are suffixes (at the ends of words). So in German, the ends of words may be much more important than in French or English.

# French letters

In French, the final letters in a lot of words are dropped when those words are spoken aloud. We expect that the final letters in French words will on average contain little information, and that is what we see. In particular, if you know the two letters that came before, then each letter beyond the first in French has little information. Even if you don't know which letters came before, then the final letters of French words have relatively little information in them.

```{r plot french letter unigrams}
plot_curves(select_language = "French")
```

```{r plot french letter bigrams}
plot_curves(select_language = "French", select_gram = "Bigram")
```

```{r french-letter-trigrams}
plot_curves(select_language = "French", select_gram = "Trigram")
```

# German letters

We expect that German, as a language with a case system which gets shown through affixation on nouns, will have relatively higher information in the final letters of words compared to English or French. We don't observe this, even in unigrams. The final letters of German words hold little information compared to the first letter of German words, which is even more noticeable when you have two letters of predictive context for each word.

```{r plot german letter unigrams}
plot_curves(select_language = "German")
```

```{r plot german letter bigrams}
plot_curves(select_language = "German", select_gram = "Bigram")
```

```{r plot german letter trigrams}
plot_curves(select_language = "German", select_gram = "Trigram")
```