### 2019-07-18 surprisal/MacWhinney poster/poster_plotting.Rmd
matched up wals features and wiki slopes

---
title: Plotting surprisal in CHILDES corpora and Switchboard English corpus
author: Josef Klafka and Dan Yurovsky 
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
# load packages
library(knitr)
library(tidyverse)
library(directlabels)
library(tidytext)
library(tidyboot)
library(here)
library(ggthemes)
library(feather)
library(janitor)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 24))
```

Get utterances
```{r get_utterances}
childes <- read_feather(here("Data/childes_words.feather")) %>%
  clean_names() %>%
  mutate(speaker_role = if_else(speaker_role == "Adult", "Parent", "Child"))

switchboard <- read_feather(here("Data/switchboard_words.feather")) %>%
  clean_names() %>%
  mutate(language = "English", speaker_role = "Adult") %>%
  select(-corpus)
```

Sample code for plotting your surprisal results.

```{r plotting}
data <- bind_rows(childes, switchboard) %>%
  filter(context %in% c(0, 2)) %>%
  mutate(language = factor(language, 
                           levels = c("English", "Spanish", 
                                      "Mandarin", "Japanese")),
         context = factor(context, levels = c(0, 2), 
                          labels = c("Unigram", "Trigram")),
         speaker_role = factor(speaker_role, 
                               levels = c("Parent", "Child", "Adult")))

label_data <- tibble(language = c("Spanish", "Spanish", "English"),
                     speaker_role = unique(data$speaker_role),
                     word_order = c(2, 4, 2),
                     s = c(6.5, 5, 6.5),
                     context = "Unigram") %>%
  mutate(context = factor(context, levels = c("Unigram", "Trigram")),
         language = factor(language, 
                           levels = c("English", "Spanish", 
                                      "Mandarin", "Japanese")))

# pdf("surprisals.pdf", width = 10, height = 5.5)
data %>%
  filter(length == 8, language != "Mandarin") %>%
   ggplot(aes(x = word_order, y = s,
                        color = speaker_role,
                        label = speaker_role)) +
  facet_grid(context~language, scales = "free") +
  geom_point(size = 1.5) +
  geom_line(size = 1.5) +
  scale_color_brewer(palette = "Set1") + 
  theme(legend.position = "none") +
  geom_text(data = label_data, size = 6) + 
  labs(x = "Word Position", y = "Average Information")
# dev.off()
```

```{r plotting features}
models <- read_feather(here("Wikipedia/models.feather"))

sf <- models %>% 
  filter(p.value == 0, estimate > 0) %>% 
  select(name) %>% 
  rename("Some Important Features" = name) %>% 
  slice(c(1, 10, 12, 15, 23, 29))

sf$Examples <- c("Noun-relative clause; relative clause-noun", 
                 "No cases; 6 cases", 
                 "SOV; VSO", 
                 "Only singular; singular and plural", 
                 "Affix; distinct word", 
                 "Prefixes; suffixes")

sf %>% grid.table()

```