### 2019-09-17 surprisal/childes_surprisal.Rmd
laplace smoothing in childes surprisal

---
title: Positional surprisal (1-3grams) in CHILDES corpora
author: Josef Klafka and Dan Yurovsky 
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
# load packages
library(knitr)
library(tidyverse)
library(directlabels)
library(childesr) #data
library(tidytext)
library(entropy)
library(tidyboot)
library(dplyr)
library(tokenizers)
library(gtools)
library(here)
library(ggthemes)
library(feather)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 16))
```

Get utterances
```{r get_utterances}
prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
zhou_utterances <- get_utterances(corpus = "ZhouDinner") #stem for pinyin
okayama_utterances <- get_utterances(corpus = "Okayama") # gloss
shiro_utterances <- get_utterances(corpus = "Shiro") #gloss
palasis_utteranecs <- get_utterances(corpus = "Palasis") #stem
wagner_utterances <- get_utterances(corpus = "Wagner") #gloss
```

```{r unigrams}
# column name must be a string, NOT just the name itself
get_unigrams <- function(utterances, speaker_role, name) {

  if (tolower(speaker_role) == "child") {
    utts <- utterances %>%
      filter(speaker_role == "Target_Child") 
  } else {
    utts <- utterances %>%
      filter(speaker_role != "Target_Child")   
  }

  utts <- utts %>%
    rename(text = name) %>%
    mutate(length = str_count(text, pattern = "[ +]+") + 1) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, text, token = stringr::str_split, pattern = "[ +]+") %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n())

  unigrams <- utts %>%
    group_by(word) %>%
    count() %>%
    ungroup() %>%
    mutate(p = n / sum(n)) %>%
    mutate(s = -log(p)) %>%
    select(word, s)

  surprisals <- utts %>%
    left_join(unigrams) %>%
     group_by(length, word_order) %>%
    summarise(mean = mean(log(s)),
              median = median(log(s)),
              se = sd(s)/sqrt(length(s))) %>%
    ungroup()
}

prov_sups <- get_unigrams(prov_utterances, "not child", "stem") %>% 
  mutate(speaker_role = "adult") %>%
  bind_rows(mutate(get_unigrams(prov_utterances, "child", "stem"), speaker_role = "child")) %>%
  mutate(language = "English")

prov_sups %>%
  filter(speaker_role == "child", length == 6, word_order == 3) %>%
  ggplot(aes(x = log(s))) + 
  geom_histogram()
```

Sample code for plotting your surprisal results.

```{r plotting}
label_data <- data_frame(length = 7, Language = "Japanese", word_order = 2,
                         speaker_role = c("Adult", "Child"), s = c(2,4))

child_surprisals %>%  
  filter(length == 12) %>%
  ggplot(aes(x = word_order, y = median)) +
    # facet_grid(Language ~ length, scales = "free") + 
    xlab("Word position") +
    ylab("Mean surprisal") +
    scale_x_continuous(breaks = 1:7) + 
    geom_line(size = 1) +
    geom_pointrange(aes(ymin = mean - 1.96 * se, ymax = mean + 1.96 * se)) + 
    theme_classic(base_size = 10) + 
    scale_color_ptol() +
    theme(legend.position = "nones") #+ 
   # geom_text(data = label_data)
```

```{r new bigrams}
adult_bigrams_prep <- adult_utterances %>% 
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  group_by(lag_word, word) %>%
  count() %>%
  filter(!is.na(lag_word)) %>%
  ungroup() %>%
  left_join(adult_unigrams, by = c("lag_word" = "word")) %>%
  mutate(cond_p = n.x / n.y) %>%
  select(lag_word, word, cond_p)

adult_bigrams <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  ungroup() %>%
  left_join(adult_unigrams) %>%
  left_join(adult_bigrams_prep) %>%
  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  group_by(length, word_order) %>%
  summarise(mean_s = mean(s)) 
```

Down here is the old code. You can run it, but you have to change the corpus name and whether you're using stem or gloss by hand.

```{r hard-coded language unigrams}
child_utterances <- prov_utterances %>%
  filter(speaker_role != "Target_Child") %>% 
  mutate(length = str_count(stem, pattern = "[ +]+") + 1) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+") %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n())

child_unigrams <- child_utterances %>%
  group_by(word) %>%
  count() %>%
  ungroup() %>%
  mutate(n = n + 1) %>%
  mutate(p = n / sum(n))

child <- child_utterances %>%
  left_join(child_unigrams, by = "word") %>%
  mutate(s = -log(p)) %>%
  ungroup() %>%
  select(length, word_order, s) %>%
  filter(length %in% c(5, 7, 9)) %>%
  group_by(length, word_order) %>%
  nest()

child %>%
  mutate(surprisal = map(data, ~tidyboot_mean(.x, s, nboot = 10))) %>%
  select(-data) %>%
  unnest(cols = c(surprisal))

  ggplot(aes(x = word_order, y = surprisal, color = as.factor(length))) + 
  geom_point() +
  geom_line()

child_surprisals <- map_dfr(1:nrow(child),
        ~bind_cols(select(slice(child, .x), length, word_order),
                               tidyboot_mean(data = unnest(slice(child, .x)),
                                             column = s, nboot = 100)))

```

```{r bigrams}
vocab_size <- child_utterances %>%
  ungroup() %>%
  distinct( word) %>% 
  nrow()

child_bigrams <- child_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  group_by(lag_word, word) %>%
  count() %>%
  filter(!is.na(lag_word)) %>%
  ungroup() %>%
  left_join(child_unigrams, by = c("lag_word" = "word")) %>%
  mutate(cond_p = (n.x + 0)  / (n.y + 0 * vocab_size)) %>%
  select(-n.y, -p)

bigram_surprisals_prep <- child_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  left_join(child_bigrams, by = c("lag_word", "word")) #%>%
  #left_join(child_unigrams, by = "word")

child <- bigram_surprisals_prep %>%
  filter(!is.na(lag_word)) %>%
  mutate(s = -log(cond_p)) %>%
 # mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  ungroup() %>%
  select(length, word_order, s) %>%
  filter(length %in% c(5, 7, 9)) %>%
  group_by(length, word_order) %>%
  nest()

child %>%
  mutate(surprisal = map(data, ~tidyboot_mean(.x, s, nboot = 10))) %>%
  select(-data) %>%
  unnest(cols = surprisal) %>%
  ggplot(aes(x = word_order, y = empirical_stat, color = as.factor(length))) + 
  geom_point() +
  geom_line()

child_bigram_surprisals <- map_dfr(1:nrow(child),
        ~bind_cols(select(slice(child, .x), length, word_order),
                               tidyboot_mean(data = unnest(slice(child, .x)),
                                             column = s, nboot = 1000)))

```

```{r trigrams}
child_trigrams <- child_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  group_by(lag_word2, lag_word1, word) %>%
  count() %>%
  filter(!is.na(lag_word1), !is.na(lag_word2)) %>%
  ungroup() %>%
  left_join(child_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
  mutate(tri_cond_p = n / n.x) %>%
  select(-n, -n.x, -cond_p)

trigram_surprisals_prep <- child_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  left_join(child_trigrams) %>% 
  left_join(select(child_bigrams, lag_word, word, cond_p), by = c("lag_word1" = "lag_word", "word")) %>% 
  left_join(child_unigrams, by = "word")

child <- trigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
                    ifelse(is.na(lag_word1), # check if not the second word
                           -log(p), # trigram if third or beyond
                           -log(cond_p)), # bigram if second word
                    -log(tri_cond_p))) %>% # unigram if first word
  ungroup() %>%
  select(length, word_order, s) %>%
  nest(s) %>%
  filter(length %in% c(5, 7, 9))

child_trigram_surprisals <- map_dfr(1:nrow(child),
        ~bind_cols(select(slice(child, .x), length, word_order),
                               tidyboot_mean(data = unnest(slice(child, .x)),
                                             column = s, nboot = 1000)))
```

```{r bootstrap data write out}
surprisals <- surprisals %>% 
  bind_rows(child_surprisals %>% mutate(context = 0) %>%
  bind_rows(child_bigram_surprisals %>% mutate(context = 1)) %>% 
  bind_rows(child_trigram_surprisals %>% mutate(context = 2)) %>% 
  mutate(speaker = "Parent", language = "French"))
surprisals %>% 
  write_feather(here("Data/all_childes_bootstrap.feather"))
```