### 2019-09-24 Paper/paper.Rmd
first couple of paras

---
title: "Speakers of diverse languages structure their utterances for efficient communication"
author: "Josef Klafka and Daniel Yurovsky"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: true
    toc: no
bibliography: ref.bib
fontsize: 11pt
link-citations: no
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
header-includes:
- \usepackage{setspace}\doublespacing
indent: true

---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(feather)
library(ggdendro)
library(gridExtra)
library(ggpmisc)

knitr::opts_chunk$set(echo = TRUE)
theme_set(theme_classic(base_size = 14))
```

```{r load data, include = F}
childes_results <- read_feather(here("Data/all_childes_bootstrap.feather"))
wiki_results <- read_feather(here("../surprisal/Data/all_spaces.feather"))
bnc_results <- read_feather(here("Data/bnc_bootstrap.feather"))
switch_results <- read_feather(here("Data/switch_bootstrap.feather"))
wiki_curves <- read_csv(here("../surprisal/Data/relative_ngrams.csv"))
wiki_lr <- read_feather(here("../surprisal/Wikipedia/models.feather"))
```

# Abstract

What role does communicative efficiency play in how we organize our utterances? In this paper, we present a novel method of examining how much information speakers in a given language communicate in each word in their utterances, surveying numerous diverse languages. We find that speakers produce frequent and informative words at regular parts of their utterances, depending on language they use, which is predictable in part from the features and genealogy of their language. This robust information distribution characterizes both spoken and written communication, and emerges in children's earliest utterances. However, in real-time communication, in-context word predictability allows listeners to process information at a constant, optimal rate, regardless of the information distribution in the language they understand.

# References