### 2019-10-28 Wikipedia/langvardata.Rmd
organized entropy branch

---
title: Wikipedia Text Mining and Entropy Analysis
author: Josef Klafka
date: `Sys.date()`
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
---

```{r setup, include=FALSE}
# load packages
library(DBI)
library(knitr)
library(tidyverse)
library(directlabels)
library(dplyr)
library(lingtypology)
library(tidytext)

knitr::opts_chunk$set(echo = TRUE)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 16))
```

```{r typology}
data <- read.csv("langvardata.csv")
data[c("language", "perimeter_log")]
```